---
title: "ch1_comm_comp"
author: "Lia Domke"
date: "10/1/2020"
output: 
  html_document: 
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include = FALSE}
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```
In chapter 1, objective 2:
2)	Determine if salmon are present in higher abundances in vegetated habitat compared to unvegetated habitat based on NOAA data from Prince of Wales (1998-2011).

Hypothesis two:
H1: Salmon species composition (including all 5 Pacific salmon species) will differ between vegetated and unvegetated habitat, with vegetated sites having more similar composition of salmon species than unvegetated sites. 

I'm interested in first, what the *overall* community composition (all species) is in the subset of habitats that we have identified in objective 2, hypothesis 2. 

First we're going to have to import the full noaa data set, and the subset of sites I've already identified. 

1. Import datasets

2. Subset full dataset by the EventIDs already identified. 

3. Data cleaning: create one dataset with species info another with environmental information

4. Data exploration

  4.1 look into transformation and standardizing for seine 
  
  4.2 transformation and standardizing for environmental
  
5. NMDS works

  5.1 Calculate dissimilarity index for both dataframes
  
  5.2 Dispersion test 
  
  5.3 Calculate NMDS
  
  4. Create species biplot
  
  5. Overlay environmental parameters?
  
6. Clean up NMDS plots

```{r libraries, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(vegan)
```


# Import data
```{r data, echo=TRUE}
noaa <- read.csv("../APECS Master repository/APECS Master repo/ALL_DATA/Lia_fish/noaa_seak_fish_atlas_CLEAN_LD.csv")

sal_env <- read.csv("Data/NOAA_SSEAK_salmon.csv", header = TRUE, stringsAsFactors = FALSE)
```

# Subset data

```{r echo=FALSE}
EventID_select <- sal_env %>%
  dplyr::select(EventID) %>%
  distinct() # create list to subset by

# subset data
noaa_sub <- subset(noaa, EventID %in% EventID_select$EventID)

# lets check it
unique(noaa_sub$EventID)
unique(noaa_sub$SiteID)

# make sure its fish
noaa_sub <- noaa_sub %>%
  filter(taxon != "invertebrate")

glimpse(noaa_sub)
```


# Data cleaning
Seperate the datasets into a species-based and environmental based dataset. Also rotate to wide form, as that will be important for transformations and standardizations
```{r message=FALSE, warning=FALSE}
# Lets seperate the datasets
# create fish only 
seine <- noaa_sub %>%
  dplyr::select(SiteID, EventID, Date, SpCode, Sp_CommonName, Sp_ScientificName, Length, Unmeasured)

# create env only 
env <- noaa_sub %>%
  dplyr::select(SiteID, EventID, Date, Season, Year, Temp, Salinity, Locale, SubLocale, Habitat, Lat1, Long1) %>%
  distinct()

# add in abundance column 
seine$abundance <- as.numeric(ifelse(is.na(seine$Length), paste(seine$Unmeasured), 1))
# sum up total abundance
seine <- seine %>%
  group_by(SiteID, EventID, Date, SpCode) %>%
  dplyr::summarise(abundance = sum(abundance))

# Fix enviromental df issue with Hab column
unique(env$Habitat) # mistake with sand-gravel
env$Habitat <- as.factor(env$Habitat)
levels(env$Habitat)[levels(env$Habitat)=="Sand-Gravel"] <- "Sand-gravel"
levels(env$Habitat)[levels(env$Habitat)=="Surfgrass"] <- "Kelp"
unique(env$Habitat)
env$Habitat <- as.character(env$Habitat)

fish_wide <- pivot_wider(seine, id_cols = c(SiteID, EventID, Date), names_from = SpCode,
                         values_from = abundance) %>%
  replace(is.na(.), 0) # add in 0s for NA cells

fish_long <- pivot_longer(fish_wide, cols = 4:76, names_to = "SpCode", values_to = "abundance")

env_wide <- pivot_wider(env, id_cols = c(SiteID, EventID, Date, Year, Locale, Lat1, Long1, Habitat), 
                      values_from = c(Temp, Salinity))

env_long <- pivot_longer(env, cols = c(Salinity, Temp),
                         names_to = "Environmental", values_to = "Value")
```

# Data exploration 
We have four dataframes to explore with. For the beach seine fish data:
fish_long
fish_wide

For environmental information (date, temp, salinity, and habitat)
env_long
env_wide
## Transformation seine
```{r ggplot theme, include=FALSE}
# set up plot themes
# Creates custom base plot theme that can adjust every graph that you use plot_theme for!

plot_theme <- function() {
  theme_bw(base_size = 14, base_family = "Avenir") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="gray96", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            strip.background = element_rect(colour = "NA", fill = "gray96"))
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```

Lets start with basic visualizations of the data
```{r figs, echo=FALSE, fig.width=7,fig.height=6,fig.cap=fig$cap("figs", "Histogram of fish abundance caught")}
par(mfrow=c(1,1))
hist(fish_long$abundance)
```

in Figure `r fig$ref("figs")` you can see that the catch data is heavily zero inflated. At many sites 0 fish of different species were caught. So lets try different transformations to shift the distribution of the data. 
A common approach is log transformations
```{r tran, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("tran", "Histogram of log transformation of total fish caught")}
par(mfrow=c(1,1))
hist(log(fish_long$abundance +1))
```
Figure `r fig$ref("tran")` shows that a log transformation does slight improve the heavy appearence of zero inflated data, however, the data is still left skewed. Note, when the data is log transformed, you have to include a constant (in this case 1) because you can't take the log of 0. 

Lets try a 4th root transformation, another common transformation
```{r tran2, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("tran2", "Histogram of log transformation of total fish caught")}
par(mfrow=c(1,1))
hist((fish_long$abundance)^ (0.25))
```
Figure `r fig$ref("tran2")` looks almost the same as a log transformation. 
Right now I'm using all the abundnaces summed together, what if we seperated by species. 
```{r sp, echo=FALSE, fig.cap=fig$cap("sp", "Histograms of fish abundances by species and site")}
g <- ggplot(fish_long, aes(x = abundance)) + geom_histogram()
g + facet_wrap(~SpCode)
```
In comparison with figure `r fig$ref("figs")` the histogram by species in figure `r fig$ref("sp")` is very similar. Lets look at it with log and 4th root transformations as well. 
```{r sp2, echo=FALSE, fig.cap=fig$cap("sp2", "Histograms of log transformed fish abundances by species and site")}
g <- ggplot(fish_long, aes(x = log(abundance+1))) + geom_histogram()
g + facet_wrap(~SpCode)
```

```{r sp3, echo=FALSE, fig.cap=fig$cap("sp3", "Histograms of fourth root transformed fish abundances by species and site")}
g <- ggplot(fish_long, aes(x = (abundance^(0.25)))) + geom_histogram()
g + facet_wrap(~SpCode)
```
Both figure `r fig$ref("sp2")` and figure `r fig$ref("sp3")` a log transformation and fourth root transformation do, help the data but it is still heavily zero inflated. 

We can test what transformation may be best using a `boxcox` test from package `MASS`
```{r}
# what does boxcox recommend? Theory suggests 4th root (for species count data)
require(MASS)
Rmod <- lm(formula = abundance+1 ~ SpCode, data = fish_long)
summary(Rmod)
boxcox(Rmod) # maybe recommends a transformation of lamba^-2 ?
```


Aside from looking at the transformation of the shape itself, we can also test the normality of the data using QQ-plots

## Standardization seine
## Transformation env
```{r env1, echo=FALSE, fig.width=7,fig.height=6,fig.cap=fig$cap("env1", "Bar graphs of salinity and temperature values from sites where seines took place")}
g <- ggplot(env_long,aes(x = reorder(as.factor(EventID), -Value), y = Value)) + geom_bar(stat = "identity")
g + facet_wrap(~Environmental)
```

```{r env2, echo=FALSE, fig.width=7,fig.height=6,fig.cap=fig$cap("env2", "Boxplot of spread of environmental values")}
ggplot(env_long) +
  geom_boxplot(aes(x = Environmental, y = Value))
```

## Standardization env


# NMDS work
## Calculate dissimilarity index for both dataframes
## Dispersion test 
## Calculate NMDS
## Create species biplot
## Overlay environmental parameters?

# Clean up NMDS plots