---
title: "chp1_obj1_nfa_3"
author: "Lia Domke"
date: "9/1/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is whats included in the dissertation as of sept 5, 2023

```{r}
sal.long.sp <- read.csv("chp1.sal.long.sp_9-1-23.csv")
```

#library
```{r}
library(tidyverse)
library(glmmTMB)
library(AICcmodavg)
library(pscl) # vuong test
library(visreg)
library(ggsignif) 
library(patchwork)
```
```{r theme settings, include=FALSE}
# Creates custom base plot theme that can adjust every graph that you use plot_theme for! First create custom plot theme

plot_theme <- function() {
  theme_bw(base_size = 24, base_family = "Helvetica") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="transparent", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank())
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```
adding global label to plot - function
```{r adding global label}
add_global_label <- function(pwobj, Xlab = NULL, Ylab = NULL, Xgap = 0.03, Ygap = 0.03, ...) {
    ylabgrob <- patchwork::plot_spacer()
    if (!is.null(Ylab)) {
        ylabgrob <- ggplot() +
            geom_text(aes(x = .5, y = .5), label = Ylab, angle = 90, ...) +
            theme_void()
    }
    if (!is.null(Xlab)) {
        xlabgrob <- ggplot() +
            geom_text(aes(x = .5, y = .5), label = Xlab, ...) +
            theme_void()
    }
    if (!is.null(Ylab) & is.null(Xlab)) {
        return((ylabgrob + patchworkGrob(pwobj)) + 
            patchwork::plot_layout(widths = 100 * c(Ygap, 1 - Ygap)))
    }
    if (is.null(Ylab) & !is.null(Xlab)) {
        return((ylabgrob + pwobj) + 
            (xlabgrob) +
            patchwork::plot_layout(heights = 100 * c(1 - Xgap, Xgap),
                                   widths = c(0, 100),
                                   design = "
                                   AB
                                   CC
                                   "
            ))
    }
    if (!is.null(Ylab) & !is.null(Xlab)) {
        return((ylabgrob + pwobj) + 
            (xlabgrob) +
            patchwork::plot_layout(heights = 100 * c(1 - Xgap, Xgap),
                                   widths = 100 * c(Ygap, 1 - Ygap),
                                   design = "
                                   AB
                                   CC
                                   "
            ))
    }
    return(pwobj)
}

```
Also - you need to create box plots of salmon catch between month and eelgrass type

Okay we've done this mulitple times I'm hoping this is the last time: 
We're comparing salmon counts by species (Chum and Pink) across habitat types and incorporating sampling timeframe as well as random effects of SiteID because we have 12 sites that we're resampled

Following Diggle et al 2002 "top-down" strategy
1. test random effects structure (this hasto be done using method = "REML")

Sal ~ Habitat + (1|Mon) + (1|SiteID)
Sal ~ Habitat + (Mon|SiteID)
Sal ~ Habitat + Mon + (1|SiteID)

Compare the model and random structure using AIC, BIC, likelihood ratio test

2. Switch to method = "ML", and determine fixed effects
We only have two potential fixed effects (month and/or habitat)
And we're truly interested in Habitat and Month is important thing we want to account for
In the end it'll include both these factors

3. Refit ideal random and fixed effect structure with method = 'REML'

We're going to use glmmTMB and see if we can get generalized mixed models for both the 
negative binomial and zeroinflated NB to work

# create each sp df
```{r}
glimpse(sal.long.sp)

chum.long <- sal.long.sp %>%
  filter(SpCode == "SALCHUM")

pink.long <- sal.long.sp %>%
  filter(SpCode == "SALPINK")
```
Also - you need to create box plots of salmon catch between month and eelgrass type

# boxplots across species and habitat and month
```{r}
p1 <- sal.long.sp %>%
   mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  filter(SpCode == "SALCHUM") %>%
  ggplot() +
  geom_boxplot(aes(x = Mon, y = abundance, fill = Habitat)) +
  plot_theme()  +
  scale_fill_manual(values = c("#339900", "#D16103"),
                    labels = c("Eelgrass", "Understory kelp"))+
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  labs(x = "Month", y = Number~of~salmon~(no.~seine^-1)) +
  theme(legend.position = "none")

p1

p2 <- sal.long.sp %>%
   mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  filter(SpCode == "SALPINK") %>%
  ggplot() +
  geom_boxplot(aes(x = Mon, y = abundance, fill = Habitat)) +
  plot_theme()  +
  scale_fill_manual(values = c("#339900", "#D16103"),
                    labels = c("Eelgrass", "Understory kelp"))+
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  labs(x = "Month", y = NULL) +
  theme(legend.position = "none")

p2

plot <- p1 + p2 + plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom")
plot

#ggsave("chp1_obj1_nfa_abund_plots.png", plot = plot, width = 14, height = 9)
```
## average by spe
```{r}
sal.long.sp %>%
  filter(SpCode == "SALCHUM" | SpCode == "SALPINK") %>%
  group_by(SpCode, Habitat) %>%
  dplyr::summarise(avg = mean(abundance),
                   sd = sd(abundance))

sal.long.sp %>%
  filter(SpCode == "SALCHUM" | SpCode == "SALPINK") %>%
  group_by(SpCode, Mon) %>%
  dplyr::summarise(avg = mean(abundance),
                   sd = sd(abundance))
```


# Chum glmmTMB
## random effects structure testing
```{r}
ch.re1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re1)

ch.re2 <- glmmTMB(abundance ~ Habitat + (1|Mon) + (1|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re2)

ch.re3 <- glmmTMB(abundance ~ Habitat + (Mon|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re3)

AICc(ch.re1); AICc(ch.re2); AICc(ch.re3)
BIC(ch.re1); BIC(ch.re2); BIC(ch.re3)
# the best random effects model is the mon as a fixed factor
```

```{r}
p.re1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re1)

p.re2 <- glmmTMB(abundance ~ Habitat + (1|Mon) + (1|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re2)

p.re3 <- glmmTMB(abundance ~ Habitat + (Mon|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re3)

AICc(p.re1); AICc(p.re2); AICc(p.re3)
BIC(p.re1); BIC(p.re2); BIC(p.re3)
```

Okay for the two species - having a random effects structure of slope rather than just random intercept structure is not possilbe 

For both species the random intercept of site is the best form using minimized AIC and BIC

2. Switch to method = "ML", and determine fixed effects
We only have two potential fixed effects (month and/or habitat)
And we're truly interested in Habitat and Month is important thing we want to account for
In the end it'll include both these factors

## determine fixed effects structure
```{r}
chum1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long,
                 REML = F, family = nbinom2(link = "log"))
summary(chum1)

# all these are significant so this is the best model? especially since I decided that doy or month wasn't actually continuous rather clustered within certain time frames
# woo here's the chum model
AICc(chum1)
BIC(chum1)
# we should test the zero inflated model tho too
```


```{r}
pink1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, REML = F, family = nbinom2(link = "log"))
summary(pink1)

pink2 <- glmmTMB(abundance ~ Habitat + Mon, data = pink.long, REML = F, family = nbinom2(link = "log"))
summary(pink2)

AICc(pink1); AICc(pink2)
BIC(pink1); BIC(pink2)
# the one without random effects is slightly better; however, we know there is structure in it so i'm not inclined to drop it.
```

## is zero inflated better?
Okay following the vignette about glmmTMB we can fit zero inflated negative binomial models using ziformula = ~1
```{r}
chum.zinf <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long, 
                 REML = F, family = nbinom2(link = "log"),
                 ziformula = ~1)
summary(chum.zinf)

chum.hurd <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long, 
                 REML = F, family = truncated_nbinom2(link = "log"),
                 ziformula = ~1)
summary(chum.hurd)

AICc(chum1); AICc(chum.zinf); AICc(chum.hurd)
BIC(chum1); BIC(chum.zinf); BIC(chum.hurd)

# vuong test compares two non-nested hypothesis test
# good way of comparing model formulas that are non nested such as zero inflated and non-zero inflated analogs
# vuong(chum1, chum.zinf)
# okay I dont think this works with random effects models, so we'll rely on AIC/BIC values
```

```{r}
pink.zinf <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, 
                 REML = F, family = nbinom2(link = "log"),
                 ziformula = ~1)
summary(pink.zinf)

pink.hurd <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, 
                 REML = F, family = truncated_nbinom2(link = "log"),
                 ziformula = ~1)
summary(pink.hurd)

AICc(pink1); AICc(pink.zinf); AICc(pink.hurd)
BIC(pink1); BIC(pink.zinf); BIC(pink.hurd)

```


For both species the normal negative binomial reduces AICc / BIC values more than the 
zero inflated negbin or the hurdle model 

3. Refit ideal random and fixed effect structure with method = 'REML'
# final models
```{r}
chum.final <- glmmTMB(abundance ~  Mon + Habitat+ (1|SiteID), 
                      data = chum.long, REML = T, family = nbinom2(link = "log"))

summary(chum.final)

library(emmeans)
emmeans(chum.final, pairwise ~  Mon, random = ~ 1|SiteID, adjust = "Tukey") 
emmeans(chum.final, pairwise ~ Habitat, random = ~1|SiteID, adjust = "Tukey")
emmeans(chum.final, pairwise ~ Habitat + Mon, random = ~1|SiteID, adjust = "Tukey")

pink.final <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID),
                      data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(pink.final)

emmeans(pink.final, pairwise ~  Mon, random = ~ 1|SiteID, adjust = "Tukey") 

```


#  model selection table
```{r}
cand.models <- list() # create an empty list to include all the candidate models
cand.models[[1]] <- ch.re1 # add in  models one at a time
cand.models[[2]] <- ch.re2

cand.models2 <- list()
cand.models2[[1]] <- chum1
cand.models2[[2]] <- chum.zinf

Modnames.re <- data.frame(names = c(paste(c(formula(ch.re1), formula(ch.re2)))),
                           method = c("REML", "REML"),
                           family = c("Neg. Binomial", "Neg. Binomial"))

Modnames.fix <- data.frame(names = c(paste(c(formula(chum1), formula(chum.zinf)))),
                           method = c("ML", "ML"),
                           family = c("Neg. Binomial", "Zero-inflated Neg. Binomial"))

# create dataframe with model names. 
cand.models

d <- data.frame()
for(i in 1:length(cand.models)) {
  mod <- cand.models[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d <- rbind(d, temp)
}

# now for the fx effects model
d2 <- data.frame()
for(i in 1:length(cand.models2)) {
  mod <- cand.models2[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d2 <- rbind(d2, temp)
}

d.chum.re <- d %>%
  left_join(Modnames.re, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.chum.fx <- d2 %>%
  mutate(family = c("Neg. Binomial", "Zero-inflated Neg. Binomial")) %>%
  left_join(Modnames.fix, by = c("model.names" = "names", "family")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.chum <- rbind(d.chum.re, d.chum.fx)
# best model includes random effect of site and habitat and month as fixed using negative binomial not zero inflated

#all.models <- as.data.frame(Modnames.all)
#write.csv(d.chum, "chp1_obj1_chum-mod-table_9-4-23.csv")
```

for pink now
```{r}
cand.models <- list() # create an empty list to include all the candidate models
cand.models[[1]] <- p.re1 # add in  models one at a time
cand.models[[2]] <- p.re2

cand.models2 <- list()
cand.models2[[1]] <- pink1
cand.models2[[2]] <- pink.zinf

Modnames.re <- data.frame(names = c(paste(c(formula(p.re1), formula(p.re2)))),
                           method = c("REML", "REML"),
                           family = c("Neg. Binomial", "Neg. Binomial"))

Modnames.fix <- data.frame(names = c(paste(c(formula(pink1), formula(pink.zinf)))),
                           method = c("ML", "ML"),
                           family = c("Neg. Binomial", "Zero-inflated Neg. Binomial"))

# create dataframe with model names. 
cand.models

d <- data.frame()
for(i in 1:length(cand.models)) {
  mod <- cand.models[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d <- rbind(d, temp)
}

# now for the fx effects model
d2 <- data.frame()
for(i in 1:length(cand.models2)) {
  mod <- cand.models2[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d2 <- rbind(d2, temp)
}

d.pink.re <- d %>%
  left_join(Modnames.re, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.pink.fx <- d2 %>%
  mutate(family = c("Neg. Binomial", "Zero-inflated Neg. Binomial")) %>%
  left_join(Modnames.fix, by = c("model.names" = "names", "family")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.pink <- rbind(d.pink.re, d.pink.fx)
#write.csv(d.pink, "chp1_obj1_pink-mod-table_9-4-23.csv")
```

# best model form table
```{r}
d.chum
summary(chum.final)
```

## visualizing chum models
using this: https://rdrr.io/cran/glmmTMB/man/predict.glmmTMB.html
```{r}
# what do we include for month
ggplot(chum.long) +
  geom_point(aes(x = Mon, y = abundance, color = Habitat)) # so the greatest catches for chum are in April and June
# lets use April for now
table(chum.long$Mon) # 135 seems reasonable

visreg(chum.final)
visreg(chum.final, scale = "response")
summary(chum.final)
# okay so for predicting random effects models likely you want to predict to the "average" site - so you might not include it in the predict fucntion. 
# chum.final includes re of siteID, habitat and month
pr.habc <- predict(chum.final, newdata = data.frame(Habitat = unique(chum.long$Habitat), Mon = "May", SiteID = NA), se.fit = TRUE)

pr.monc <- predict(chum.final, newdata = data.frame(Habitat = "Eelgrass", Mon = unique(chum.long$Mon), SiteID = NA), se.fit = TRUE)

pred <- exp(pr.habc$fit) 
lwr <- exp(pr.habc$fit - (1.96*pr.habc$se.fit))
upr <- exp(pr.habc$fit + (1.96*pr.habc$se.fit))

pred.m <- exp(pr.monc$fit)
lwr.m <- exp(pr.monc$fit - (1.96*pr.monc$se.fit))
upr.m <- exp(pr.monc$fit + (1.96*pr.monc$se.fit))

chum.long$Habitat <- as.character(chum.long$Habitat)

pr.habc.df <- data.frame(cbind(pred, lwr, upr, unique(chum.long$Habitat))) %>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(pred),
         lwr = as.numeric(lwr),
         upr = as.numeric(upr))

chum.long$Mon <- as.character(chum.long$Mon)

pr.mon.df <- data.frame(cbind(pred.m, lwr.m, upr.m, unique(chum.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(pred.m),
         lwr = as.numeric(lwr.m),
         upr = as.numeric(upr.m))
```
Now put it together
```{r}
chum.long$Mon<- plyr::revalue(chum.long$Mon, c("Apr"="April", "May"="May", "Jun"= "June", "Jul"="July"))
                           
# make sure when you're putting the graph together that you include lwr and upr without the fit cause those are the lwr and upr confidence intervals. 
habc <- ggplot(pr.habc.df) +
  geom_col(aes(x = Habitat, y = fit, fill = Habitat)) + #fill = "#7c7c7c") + #fill = "lightgray") +
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = chum.long, aes(x = Habitat, y = abundance)) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans = "log10") +
  scale_fill_manual(values = c("#339900", "#D16103")) +
  theme(legend.position="none") +
  labs(y = Juvenile~Chum~Salmon~(no.~seine^-1)) #+
  #geom_signif(data = chum.long, aes(x = Habitat, y = abundance), 
   #                    comparisons = list(c("Eelgrass", "Kelp")),
    #          map_signif_level=TRUE, textsize = 8, annotations = c("p-value = 0.062"))
habc

summary(chum.final)
# now doy
monc <- pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "#7c7c7c") + #fill = "lightgray") +
  geom_errorbar(aes(x = Mon, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = chum.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans="log10") +
  labs(x = "Month", y = NULL) #+
  #geom_signif(data = chum.long, aes(x = Mon, y = abundance), 
   #           comparisons= list(c("April", "May", "June","July")),
    #          map_signif_level=T, manuel = T)


#                       comparisons = list(c("April", "May", "June", "July")),
 #             map_signif_level=TRUE, textsize = 8)
monc
summary(chum.final)
 glmm.chum.plot <- habc + monc + plot_annotation(tag_levels="a")

glmm.chum.plot.talk <- monc + habc
glmm.chum.plot
#ggsave("chp1_obj1_glmm.chum.plot_colored.png", plot = glmm.chum.plot.talk, width = 16, height = 7)
#ggsave("chp1_obj1_glmm.chum.plot.png", plot = glmm.chum.plot, width = 14, height = 8)
```

# best pink model
```{r}
d.pink
summary(pink.final)
```

## visualziing pink model
```{r}
# what do we include for month
ggplot(pink.long) +
  geom_point(aes(x = Mon, y = abundance, color = Habitat)) # so the greatest catches for pink are in April and May
# lets use May for now
table(pink.long$Mon) # 135 seems reasonable
table(pink.long$Mon, pink.long$abundance>0) # more non-zero catches in may compared to any other month... could this be way the CI doesn't go as low? 

visreg(pink.final)
visreg(pink.final, scale = "response")
summary(pink.final)
# okay so for predicting random effects models likely you want to predict to the "average" site - so you might not include it in the predict fucntion. 
# chum.final includes re of siteID, habitat and month
pr.habp <- predict(pink.final, newdata = data.frame(Habitat = unique(pink.long$Habitat), Mon = "May", SiteID = NA), se.fit = TRUE)

pr.monp <- predict(pink.final, newdata = data.frame(Habitat = "Eelgrass", Mon = unique(pink.long$Mon), SiteID = NA), se.fit = TRUE)

predict(pink.final, newdata = data.frame(Habitat = "Eelgrass", 
                                         Mon = unique(pink.long$Mon), 
                                         SiteID = NA), 
        se.fit = TRUE, type = "response")

# if 
#log_normal_correct_p<- (sd(residuals(pink.final))^2)/2


predp <- exp(pr.habp$fit)
lwrp <- exp(pr.habp$fit - (1.96*pr.habp$se.fit))
uprp <- exp(pr.habp$fit + (1.96*pr.habp$se.fit))

predp.m <- exp(pr.monp$fit)
lwrp.m <- exp(pr.monp$fit - (1.96*pr.monp$se.fit))
uprp.m <- exp(pr.monp$fit + (1.96*pr.monp$se.fit))

pink.long$Habitat <- as.character(pink.long$Habitat)

pr.habp.df <- data.frame(cbind(predp, lwrp, uprp, unique(pink.long$Habitat)))%>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(predp),
         lwr = as.numeric(lwrp),
         upr = as.numeric(uprp))

pink.long$Mon <- as.character(pink.long$Mon)

pr.monp.df <- data.frame(cbind(predp.m, lwrp.m, uprp.m, unique(pink.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(predp.m),
         lwr = as.numeric(lwrp.m),
         upr = as.numeric(uprp.m))
```

```{r}
pink.long$Mon <- plyr::revalue(pink.long$Mon, c("Apr"="April", "May"="May", "Jun"= "June", "Jul"="July"))
                           

habp <- ggplot(pr.habp.df) +
  geom_col(aes(x = Habitat, y = fit, fill = Habitat)) + #, fill = "lightgray") +
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = pink.long, aes(x = Habitat, y = abundance)) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans = "log10") +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
   #                  breaks = c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000)) +
  scale_fill_manual(values = c("#339900", "#D16103")) +
  theme(legend.position="none") +
  labs(y = Juvenile~Pink~Salmon~(no.~seine^-1))
habp


# now doy
monp <- pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "#7c7c7c") + # fill = "lightgray"
  geom_errorbar(aes(x = Mon, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = pink.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  #ylim(c(0, 8000)) +
  plot_theme() +
  #ylim(c(0, 9000)) +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
   #                  breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) +
  labs(x = "Month", y = NULL) 
monp

glmm.pink.plot <- habp + monp + plot_annotation(tag_levels="a")
glmm.pink.plot.colored <- monp + habp

p_ranges_y <- c(ggplot_build(glmm.pink.plot[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(glmm.pink.plot[[2]])$layout$panel_scales_y[[1]]$range$range)

glmm.pink.plot2 <- glmm.pink.plot &
  ylim(min(p_ranges_y), max(p_ranges_y)) &
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) 


p_ranges_y2 <- c(ggplot_build(glmm.pink.plot.colored[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(glmm.pink.plot.colored[[2]])$layout$panel_scales_y[[1]]$range$range)

glmm.pink.plot.colored2 <- glmm.pink.plot.colored &
  ylim(min(p_ranges_y2), max(p_ranges_y2)) &
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) 

#ggsave("chp1_obj1_glmm.pink.plt_logscale.colored.png", plot = glmm.pink.plot.colored2, width = 16, height = 7)

#ggsave("chp1_obj1_glmm.pink.plot_logscale.png", plot = glmm.pink.plot2, width = 14, height = 8)
```

Non-exponentiated CI of the predicted pink.final
```{r}
pr.habp <- predict(pink.final, newdata = data.frame(Habitat = unique(pink.long$Habitat), Mon = "May", SiteID = NA), se.fit = TRUE)

pr.monp <- predict(pink.final, newdata = data.frame(Habitat = "Eelgrass", Mon = unique(pink.long$Mon), SiteID = NA), se.fit = TRUE)

predp <- pr.habp$fit
lwrp <- pr.habp$fit - (1.96*pr.habp$se.fit)
uprp <- pr.habp$fit + (1.96*pr.habp$se.fit)

predp.m <- pr.monp$fit
lwrp.m <- pr.monp$fit - (1.96*pr.monp$se.fit)
uprp.m <- pr.monp$fit + (1.96*pr.monp$se.fit)

pink.long$Habitat <- as.character(pink.long$Habitat)

pr.habp.df <- data.frame(cbind(predp, lwrp, uprp, unique(pink.long$Habitat)))%>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(predp),
         lwr = as.numeric(lwrp),
         upr = as.numeric(uprp))

pink.long$Mon <- as.character(pink.long$Mon)

pr.monp.df <- data.frame(cbind(predp.m, lwrp.m, uprp.m, unique(pink.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(predp.m),
         lwr = as.numeric(lwrp.m),
         upr = as.numeric(uprp.m))

ggplot(pr.habp.df) +
  geom_col(data = pr.habp.df, aes(x = Habitat, y = fit, fill = Habitat)) + #fill = "lightgray") +
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2, data = pr.habp.df) #+
  #geom_jitter(data = pink.long, aes(x = Habitat, y = log(abundance+1))) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans = "log10") +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10)) +
  #scale_fill_manual(values = c("#339900", "#D16103")) +
  theme(legend.position="none") +
  labs(y = Juvenile~Pink~Salmon~(no.~seine^-1))

pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "#7c7c7c") +
  geom_errorbar(aes(x = Mon, ymax = fit + upr, ymin = fit - lwr), width = 0.2) +
  #geom_jitter(data = pink.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  #ylim(c(0, 8000)) +
  plot_theme() +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10)) +
  #ylim(c(0, 10000)) +
  labs(x = "Month", y = NULL) 
```


# best model tables
```{r}
library(sjPlot)
summary(pink.final)
summary(chum.final)
tab_model(chum.final, pink.final,
          transform = NULL,
          dv.labels = c("Juvenile Chum Salmon", "Juvenile Pink Salmon"),
          show.est = TRUE, string.est = "Estimate", show.r2 = FALSE,
          show.stat = TRUE, string.stat = "Statistic",
          p.threshold = c(0.1, 0.05, 0.01, 0.001),
          show.se = TRUE, show.ci = FALSE, string.p = "P-Value",
          string.se = "Standard Error", p.style = "numeric")
           
          
          
          # pred.labels = c("Intercept", "Habitat [kelp]","Non-linear date", 
          #                  "Site (random)", "Date"))
```

