---
title: "ch1_mixed_models"
author: "Lia Domke"
date: "10/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Mixed-effects models
The main  hypothesis is that structured habitats or vegetated habitats are important structure for salmon in the nearshore. To test that using the NOAA dataset, I ran some simple glm and gam models by species to look at the main effect of habitat. However, in those model, I was not able to account for the sampling design that actually pairs habitat type within site. 

Based on Schielzeth & Nakagawa (2012), I've determined that my main effect factor 1 is Habitat type and my main effect factor 2 is site. At 6 sites, I have a fully crossed sampling design where every habitat type (bedrock, kelp, and eelgrass) were sampled at each site. 

By using mixed-effect model as a modelling framework with a fully crossed sampling design--this allows partitioning of variance between the four main sources of variances:

1. Main effect variance explained by factor 1

2. Main effect variance explained by factor 2

3. Interaction variance -- variance in the response explainby combination of site x habitat **after controlling** for the average effect of the habitat acoss all sites and the average effect of the site on habitat type. 

4. Residual variance -- variance in the response unexplained by the habitat, site, and their interaction. If you *exclude* or *remove* the interaction term then the variance is POOLED in residual variance.  

Important decision making points when fitting a mixed model. Determine which factors are fixed versus random is important. Including random factors should not be done lightly as it completely changes the interpretation of the model estimates. See the blog post and thread for more information on the difference between the two and when to use random effects [link](https://dynamicecology.wordpress.com/2015/11/04/is-it-a-fixed-or-random-effect/). 

Based on advise from Franz's course--when fitting mixed effect models and obtaining parameter estimations there are two main methods, Maximum Likelihood (ML) and Restricted Maximum Likelihood (REML). You **cannot** use REML to compare models with different fixed effects structure, rather ML is the best use for comparison of different models. After a "best" model is determiend, parameter estimation/CI using REML should be used. 

Model set up:
We selected 6 sites, with three habitat types that were each sampled once for salmon CPUE.

Questions:
1. What is the abundance of salmon in a typical habitat type
2. What is the variation of salmon among habitats (among group variance)
3. What is the variation of salmon within sites (within group variance)

yij = beta + bi + Eij

beta - mean abundance across habitats
bi - random variable representing the deviation from the population mean
bi ~ N(0, sigma-b^2)
Eij - random variable representing the deviation from observation j from the mean density for site 

## Assumptions
residuals are assumed to be independent, normally distributed random variables with **constant variance**

random effects are assumed to be **independent** and normally distributed

## Model packages options
nlme, functions : lme, nlme    -- can accommodate residual correlations

lme4 - more stable algorithms, can fit generalized mixed effects models but does not accommodate residual correlation, functions: lmer, nlmer, glmer

glmm - gmm using monte carlo likelihood estimation

```{r library, message=FALSE, warning=FALSE, include=FALSE}
library(nlme)
library(ggplot2)
library(lme4)
library(emmeans)
library(tidyverse)
```


```{r data, include=FALSE}
# Need to import data from another Rmarkdown where I cleaned the data and subset it by species. 
# File name: chp1_sal_and_veg(noaa).Rmd in Project Chp1_salmon
pink <- read.csv("Data/Noaa_pink.csv", stringsAsFactors = FALSE, header = TRUE)
```

```{r theme settings, include=FALSE}
# Creates custom base plot theme that can adjust every graph that you use plot_theme for!

plot_theme <- function() {
  theme_bw(base_size = 14, base_family = "Avenir") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="gray96", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            strip.background = element_rect(colour = "NA", fill = "gray96"))
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```

```{r fig labeling, include = FALSE}
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

```{r data visualizations}
# quick look at the data were using
str(pink)

table(pink$site_pair) # observations at each site pair. 
```

```{r by.site, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("by.site", "Pink salmon abundance per seine in different habitats at nine different sites")}
p1 <- ggplot(pink, aes(x = Habitat, y = abundance)) + 
  geom_bar(stat = "identity") + plot_theme() + 
  ylab("Abundance")
p1 + facet_wrap(~site_pair) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

In `r fig$ref("by.site")` you can see that there are variable numbers of pink salmon caught at each site. Clearly the graph is dominated by the catch of almost 800 pink salmon at Moira Sound. In order to get a better idea, I'll free up the y scale. 

```{r by.site.free, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("by.site.free", "Pink salmon abundance per seine in different habitats at nine different sites with free y axis scale")}
p2 <- ggplot(pink, aes(x = Habitat, y = abundance)) + 
  geom_bar(stat = "identity") + plot_theme() + 
  ylab("Abundance")
p2 + facet_wrap(~site_pair, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

`r fig$ref("by.site.free")` shows that if you look at relative abundances at each site between habitat types, at 4 of the 9 sites more salmon were caught in bedrock over the other three habitats. Kelp and eelgrass had some salmon catch at different sites. However, clearly both sites and habitat are important is determining highly variable salmon abundances. 

Want to use mixed models to account for site when looking at the influence of habitat type. 

```{r hist, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("hist", "Histogram of the raw frequency of pink salmon and log transformed pink salmon caught overall at all sites and habitats")}
# reminder that pink abundance is not normal and we should be using some kind of transformation - log is better, but honestly not the greatest
par(mfrow=c(1,2))
hist(pink$abundance)
hist(log(pink$abundance+1))
```

Likely need some sort of transformation to approximate normality, however, how strict is the normality assumption when implementing mixed models? 
Lets subset the data so that we're looking at sites with complete observations to have a fully crossed design. 
```{r include=FALSE}
# lets subset by the sites with complete observations to have a fully crossed design
unique(pink$site_pair)
pink.sub <- pink %>%
  filter(site_pair != "Klawock Inlet") %>%
  filter(site_pair != "Ballena Island") %>%
  filter(site_pair != "South Wadleigh")

hist(log(pink.sub$abundance+1))
```

```{r echo=FALSE}
par(mfrow=c(1,2))
boxplot(abundance ~ site_pair, data = pink.sub)
boxplot(abundance ~ Habitat, data = pink.sub) # in pink sub no sand/gravel sites
```

Lets look at some different ways to write mixed models. 
```{r}
mxd <- lmer(abundance ~ Habitat + (1 | site_pair), data = pink.sub)
# model doesn't fit well. because the fitted model is near singularity, however, I'm not sure what that means. 
# indicates that the variances of one or more of the linear combinations of effects are close to (or actually) zero. 
summary(mxd)
```

```{r}
# lets try a diff package - lmList
# This fits the same fixed effect model **seperately** for each group (site?)
lmList_mxd <- lmList(abundance ~ Habitat | site_pair, data = pink.sub)
# Examin the parameter estimates (the coefficients from multiple linear regression) and the standard errors, t-statistic and p-value are testing the null hypothesis that the coefficients are equal to zero. 
summary(lmList_mxd) # note that there doesn't seem to be enough info the generate a ttest

par(mfrow=c(1,1))
a<-coef(lmList_mxd)$"(Intercept)"
hist(a) # the intercept parameters should be appox normally distributed in order to model site specific effects as random effects. This graph looks like its right skewed
qqnorm(a); qqline(a) # some spread, is this dataset too small to do this kind of model? 

plot(lmList_mxd) # residuals? 
```

Based on Franz's lectures from FISH 604, I've also tried using `nlme::lme`
```{r}
# First I'm going to fit the model using habitat as the fixed effect and site as the random. Because I'm only interested in this model and won't be doing model comparisons with different models, I'm going to fit the model using REML rather than ML. 
mxd_pink <- lme(abundance ~ Habitat, data = pink, random = ~ 1 | site_pair,
method="REML")
```
Lets take a look at the summary table for this model. 
```{r}
summary(mxd_pink)
```

Based on my conversations with Nicole, she said that the relative comparison of the random effect stdDev of intercept and residual are important. In her case the intercept residual was almost the same as the otterID (random) residual indicating that the OtterID plays almost the same role in determining carbon as season. 

In my case, the residual of the random effect is substantially larger than the intercept (main effects?). So I take that to mean that there is a lot more residual accounted for by siteID rather than Habitat. Indicating that siteID is more important in determining pink salmon abundance. However, I'm not positive. 

If we disregard this and continue to try and interpret the results. When looking at the fixed effects the mean value of bedrock (indicated by intercept) is substantially greater than the mean values in the other habitats and appears significantly different. Might be worthwhile to look into pairwise comparisons potentially using tukey's test. 

My other question is about this: Because this is a mixed model, aren't you accounting for the variation in siteID in order to look at the effect of habitat? Does it really matter that the random effect residual is so high? And Where is the residual variation reported? 

Lets test the role of habitat using an anova

```{r}
av <- anova(mxd_pink)
av
```
```{r include=FALSE}
pv <- av$`p-value`[2]
```
What does it mean that intercept is significant? 
At alpha 0.1, the group habitat is significant (p-value `r pv`)

Lets look at the significants between habitats
```{r}
tv <- lsmeans(mxd_pink, pairwise~Habitat, adjust = "tukey")
tv
```

Although it does look liek the role of habitat can be important there are no significant differences between habitats for pink salmon abundances. 

```{r}
plot(tv$lsmeans)
```

Okay, so thats good to know. But we subset the data we had for a fully crossed design. What if we use the full dataset and use Hab_type to look just vegetated and unvegetated habitats? It will still be a fully crossed design but will define the habitats as vegetated or not. 
```{r}
mxd_v.unv <- lme(abundance ~ Hab_type, data = pink,
                     random = ~1 | site_pair, method = "REML")
```
```{r}
summary(mxd_v.unv)
```
Well similar story here, the random effects residuals are still substantially larger than the intercept--which is questionable. 

Lets look at an anova
```{r}
av2 <- anova(mxd_v.unv)
av2
```
```{r include=FALSE}
pv2 <- av2$`p-value`[2]
```
Habitat is significant at alpha = 0.1 (p-value `r pv2`)

Lets look at the significants between habitats
```{r}
tv2 <- lsmeans(mxd_v.unv, pairwise~Hab_type, adjust = "tukey")
tv2
```

Oh duh. Its the same I think. 
```{r}
plot(tv2)
```

CI overlab for different habitat types veg and unvegetated. 
