---
title: "phyllospadix_biomass_regression"
author: "Lia Domke"
date: "3/8/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We're curious to see if there is a relationship between *longest* leaf length to *total* shoot biomass FOR PHYLLOSPADIX. We know there is a relationship with Zostera marina. 

Data available: In 2021 sites were surveyed for seagrass and some had phyllosapdix. At each site, 5-0.25m^2 quadrats were surveyed and 15 entire shoots (i.e. multiple leaves) were removed and brought back to the lab to be measured (in cm) and weighed (in g). 

We are interested in the length of the longest leaf and the *WET* biomass. (for Zostera marina we looked at dry biomass)

If there is a strong relationship, then I would like to use the linear model from 2021 to predict biomass values based on longest leaf length in 2019 and 2021.

read in data
- this data comes from the cleaned data from the APECS_DM files
```{r}
wgt <- read.csv("../APECS_DM/Data/Habitat_data/hab_weight_2021_clean_10-11-22.csv", stringsAsFactors=F, header = T)
lng <- read.csv("../APECS_DM/Data/Habitat_data/hab_length_2021_clean_10-11-22.csv", stringsAsFactors=F, header =T)
sp.names <- read.csv("../APECS_DM/Data/Habitat_data/hab_species_name_2021_9-27-22.csv", stringsAsFactors=F, header =T)
qrt <- read.csv("../APECS_DM/Data/Habitat_data/hab_quadrat_2019_clean_10-11-22.csv", stringsAsFactors=F, header =T)


qrt22 <- read.csv("../APECS_DM/Data/Habitat_data/hab_quadrat_2022_clean_3-14-23.csv", stringsAsFactors=F, header =T)
lng22 <- read.csv("Data/hab_length_2022_Zosbiomass_10-18-23.csv", stringsAsFactors= F, head = T)
wgt22 <- read.csv("../APECS_DM/Data/Habitat_data/hab_weight_2022_clean_3-14-23.csv", stringsAsFactors= F, header =T)

```

libraries
```{r}
library(tidyverse)
```


need to find instances where there is length and weight information for a Phyllo plant
```{r}
head(wgt)
unique(wgt$species)

wgt %>%
  filter(grepl("Phyllospadix", species)) # this has the total weight for all phyllospadix plants collect (in total) by quadrat,
# i think this isn't helpful

lng22.phyll<- lng22 %>%
  filter(grepl("Phyllospadix", species)) # all instances have lengths but no weights in 2022 

indiv <- lng %>%
  filter(grepl("Phyllospadix", species),
         !is.na(weight_g))

# how many quadrats?
unique(indiv$quadrat_ID) # 8 quadrats for a total of 131 observations
```

what is the relationship?
```{r}
plot(indiv$weight_g, indiv$length_mm) # looks like there's somewhat of a relationship

fit.lm <- lm(weight_g ~ length_mm, data = indiv)
summary(fit.lm) # length is sig diff than 0 (parameters), adjusted rsquared is 0.75 pretty good
par(mfrow = c(2,2))
plot(fit.lm, which = 1:4) # the qq plot looks pretty good, the residuals look like the spread does increase some across the fitted values. potential outliers at 50 and 55, with 50 looking parituclarily influenctial. 

AIC(fit.lm)

e <- residuals(fit.lm)
shapiro.test(e)

library(e1071)
skewness(indiv$weight_g)
skewness(indiv$length_mm)

# Does a boxcox indicate a transformation is necessary? 
library(MASS)
boxcox(fit.lm)
boxcox(fit.lm, lambda=seq(from=0, to=0.8, by=.01)) # lambda around 0.4

# log log tranformations
plot(log(indiv$weight_g) ~ log(indiv$length_mm))
fit.lmlog <- lm(log(weight_g) ~ log(length_mm), data = indiv)
summary(fit.lmlog)
plot(fit.lmlog) # fitted patterns look better

AIC(fit.lm, fit.lmlog) # log - log transformation is better
```

fitted model:
fit.lmlog

predict values based on this relationshipo

```{r}
summary(fit.lmlog)
# formula :
#   log (y) = beta0 + beta1 * log (x) 
#   log (dw) = -7.78 + 1.30 * log (max_length) 
# create a new dataframe with just the leaf lengths from 2021

phyll.lng <- lng %>%
  filter(grepl("Phyllospadix", species),
         is.na(weight_g)) # all lengths no weight

lng22.phyll %>%
  filter(is.na(weight_g)) # its the same as above, we're good

newx <- data.frame(length_mm = as.numeric(phyll.lng$length_mm))
newx22 <- data.frame(length_mm = as.numeric(lng22.phyll$length_mm))

# use the log-log linear model (equation above) to calculate the dry mass of the shoot with 95% CI 
# make sure to exp() the values because its a log-log relationship
pr.lm <- data.frame(predict(fit.lmlog, newdata = newx, interval = "confidence", level = 0.95))

# okay because we had to log-transform the data in order to make it fit 
# normality it requires a log-normal transformation which is equal to the 
# squared(standard deviation of the residuals)/2, so variance/2
correction <- (sd(fit.lmlog$residuals)^2)/2 # this is the correction factor
corr.fit <- exp(pr.lm$fit + correction) # you add it to the fitted LOG SPACE and then exponentiate it
corr.lwr <- exp(pr.lm$lwr + correction) # also needs to be done to the confidence intervals
corr.upr <- exp(pr.lm$upr + correction)

hist(corr.fit)
hist(exp(pr.lm$fit))
range(corr.fit) 
range(indiv$weight_g)

pr.lm22 <- data.frame(predict(fit.lmlog, newdata = newx22, interval = "confidence", level = 0.95))
corr.fit22 <- exp(pr.lm22$fit + correction)
corr.lwr22 <- exp(pr.lm22$lwr + correction)
corr.upr22 <- exp(pr.lm22$upr + correction)


# graph these new predicted values
# create data.frame first
newdata <- cbind(phyll.lng, corr.fit)
ggplot() + geom_point(data = newdata, aes(x = length_mm, y = corr.fit), color = "green") + 
  theme_classic() + 
  geom_point(data = indiv, aes(x = length_mm, y = weight_g), color = "blue") +
  geom_smooth(method = "lm")

newdata22 <- lng22.phyll %>%
  dplyr::select(-c(fit, lwr, upr, X, X.1)) %>%
  cbind(corr.fit22) %>%
  cbind(corr.lwr22) %>%
  cbind(corr.upr22)
  
ggplot() + 
  geom_ribbon(data = newdata22, aes(x = length_mm, ymin = corr.lwr22, ymax = corr.upr22), fill = "gray")+
  geom_point(data = indiv, aes(x = length_mm, y = weight_g), color = "blue") +
  geom_point(data = newdata22, aes(x = length_mm, y = corr.fit22))
  
```

Estimate biomass for length of blades
```{r}
str(newdata)
str(qrt)


# what quads have pyllos
lng %>%
  filter(grepl("Phyllospadix", species)) # all habs including measured and unmeasured, 188 observations

# what is the range of values for weight - does it make sense?
range(newdata$corr.fit)
range(indiv$weight_g) # yea more or less they match, lets make the new data the weight data now. 

head(newdata)
phyll.weight <-newdata %>%
  mutate(weight_g = corr.fit) %>%
  dplyr::select(-c(corr.fit)) # 57 observations now weight

lng.measured <- lng %>%
  anti_join(phyll.weight, by = c("plant_ID")) %>% # lng was 2123; removed the unmeasured phyll = 2066
  rbind(phyll.weight) # add back in the rows with measured values

# check to make sure it worked
lng.measured %>%
  filter(grepl("Phyllospadix", species),
         is.na(weight_g)) # none unmeasured!


## now for 2022
lng22 %>%
  filter(grepl("Phyllospadix", species))

range(newdata22$corr.fit22)
range(indiv$weight_g)

newdata22 <- newdata22 %>%
  dplyr::rename(fit = corr.fit22, lwr = corr.lwr22, upr = corr.upr22)

lng22.red <- lng22 %>%
  anti_join(newdata22, by = c("plant_ID", "quadrat_ID", "species", "length_mm", 
                              "weight_g", "notes", "length_cm")) %>%
  dplyr::select(-c(X.1, X))

lng22.phyll.weight <- rbind(lng22.red, newdata22)


# we want to keep the upr an dlwr for the mean weight based on length 
#write.csv(lng22.phyll.weight, "../APECS_DM/Data/Habitat_data/hab_length_2022_clean_biomass_10-9-23.csv")
#write.csv(lng.measured, "../APECS_DM/Data/Habitat_data/hab_length_2021_measuredphyllo_10-9-23.csv")
```