---
title: "SalmonByRegion_NOAA"
author: "Lia Domke"
date: "10/9/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Interested in looking at salmon species by region

Steps:

1. Load Libraries

2. Load Data, fish by region

3. Filter by salmon only

4. Add back in zeros where no salmon caught by region

# Libraries
```{r libraries, include=FALSE}
library(tidyverse)
library(cowplot)
```

# Data
```{r data}
seni <- read.csv("Data/noaa_region_seni.csv", stringsAsFactors = FALSE, header = TRUE)
seno <- read.csv("Data/noaa_region_seno.csv", stringsAsFactors = FALSE, header = TRUE)
seso <- read.csv("Data/noaa_region_seso.csv", stringsAsFactors = FALSE, header = TRUE)
sesi <- read.csv("Data/noaa_region_sesi.csv", stringsAsFactors = FALSE, header = TRUE)
```

District column
```{r include=FALSE}
seni$district <- "seni"
seno$district <- "seno"
seso$district <- "seso"
sesi$district <- "sesi"

# join together
noaa <- rbind(seni, seno) %>%
  rbind(seso) %>%
  rbind(sesi) %>%
  filter(taxon != "invertebrate")

# fix hab levels 
noaa$Habitat <- as.factor(noaa$Habitat)
levels(noaa$Habitat)[levels(noaa$Habitat)=="Sand-Gravel"] <- "Sand-gravel"
levels(noaa$Habitat)[levels(noaa$Habitat)=="Surfgrass"] <- "Kelp"

# Add in veg/unveg
noaa <- noaa %>%
  mutate(Hab_type = ifelse(Habitat %in% c("Bedrock", "Sand-gravel"), "Unvegetated", "Vegetated"))
```

```{r include=FALSE}
# create environmental/site info
noaa_info <- noaa %>%
  dplyr::select(SiteID, EventID, Date, Year, Mon, Gear, Temp, Salinity, Region, Locale, Location, SubLocale, Hab_type,
                Nickname, Habitat, Lat1, Long1, taxon, district) %>%
  distinct() %>%
  dplyr::rename(Latitude = Lat1, Longitude = Long1)

```

# Filter by salmon species
```{r include=FALSE}
noaa$abundance <- as.numeric(ifelse(is.na(noaa$Length), paste(noaa$Unmeasured), 1))

noaa_sal <- noaa %>%
  filter(SpCode == "SALPINK" |
           SpCode == "SALCHUM" |
           SpCode == "SALCHIN" |
           SpCode == "SALCOHO" |
           SpCode == "SALSOCK") %>%
  group_by(SiteID, EventID, Date, Habitat, SpCode) %>%
  dplyr::summarise(abundance = sum(abundance)) 

```

# Add zeros 
Okay so that summarised by spcode and eventid how many salmon *were* caught, but need to add back in the zeros too. 

```{r include=FALSE}
# add back in env info
sal <- full_join(noaa_sal, noaa_info)

df_wide <- pivot_wider(sal, id_cols = c(SiteID, EventID, Locale, SubLocale, Date, Mon, Year, Latitude, Longitude, Habitat, Hab_type, district), names_from = SpCode, values_from = abundance) %>%
  dplyr::select(-"NA") %>%
  replace(is.na(.), 0)

# wide -- > long, convserves zeros and site/event ID information

sal_long <- pivot_longer(df_wide, cols = c(SALCHIN, SALCOHO, SALPINK, SALCHUM, SALSOCK), names_to = "SpCode", values_to = "abundance")
```

# Visualize data
```{r, include = FALSE}
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```
```{r theme settings, include = FALSE}
# Creates custom base plot theme that can adjust every graph that you use plot_theme for!

plot_theme <- function() {
  theme_bw(base_size = 14, base_family = "Avenir") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="gray96", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            strip.background = element_rect(colour = "NA", fill = "gray96"))
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```

First lets look at total salmon abundances by species in different months
```{r by.month, echo = FALSE, fig.width=7, fig.height=6, fig.cap=fig$cap("by.month", "Total number of salmon caught by month overall in Southeast Alaska"), message=FALSE, warning=FALSE}
sal_long$Mon <- as.factor(sal_long$Mon)

sal_long %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Jan","Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep"))) %>%
  group_by(SpCode, Mon) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() +
  geom_bar(aes(x = Mon, y = total, fill = SpCode), stat="identity") +
  plot_theme() + scale_fill_brewer(name = "Species", labels = c("Chinock", "Chum", "Coho", "Pink", "Sockeye"), palette = "Dark2")
  
```


Now, I divided total number of salmon caught by different districts in Southeast Alaska
```{r by.district, echo = FALSE, fig.width=10, fig.height=6, fig.cap=fig$cap("by.district", "Total number of salmon caught by month and by district Southeast Alaska"), message=FALSE, warning=FALSE}

g1 <- sal_long %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Jan","Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep"))) %>%
  group_by(SpCode, Mon, district) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() +
  geom_bar(aes(x = Mon, y = total, fill = SpCode), stat="identity") +
  plot_theme()
g1 + facet_wrap(~district) + scale_fill_brewer(name = "Species", 
                                               labels = c("Chinock", "Chum", "Coho", "Pink", "Sockeye"),
                                               palette = "Dark2")
```

Now lets go district by district and look at salmon numbers for each species. 

In addition we'll drop the months that we don't care about. Including only April, May, June, and July

# First SENI (Southeast Northern Inside - Juneau/Lynn Canal Area)
```{r seni.hab, echo = FALSE, fig.width=10, fig.height=6, fig.cap=fig$cap("seni.hab", "Number of salmon caught in summer months at four different habitat types in Southeast Northern Inside area with free y axis scale"), message=FALSE, warning=FALSE}
seni1 <- sal_long %>%
  filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  filter(district == "seni") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(SpCode, Mon, Habitat) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity") +
  plot_theme() 
seni1 + facet_wrap(~SpCode, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

# Next SENO (Southeast Northern Outside - Sitka/Outside Chichagof)
```{r seno.hab, echo = FALSE, fig.width=10, fig.height=6, fig.cap=fig$cap("seno.hab", "Number of salmon caught in summer months at four different habitat types in Southeast Northern Outside area with free y axis scale"), message=FALSE, warning=FALSE}
seno1 <- sal_long %>%
  filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  filter(district == "seno") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(SpCode, Mon, Habitat) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity") +
  plot_theme() 
seno1 + facet_wrap(~SpCode, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

# Next SESI (Southeast Southern Inside - Protected western POW/northern KTN up to southern Fredrick Sound)

```{r sesi.hab, echo = FALSE, fig.width=10, fig.height=6, fig.cap=fig$cap("sesi.hab", "Number of salmon caught in summer months at four different habitat types in Southeast Southern Inside area with free y axis scale"), message=FALSE, warning=FALSE}
sesi1 <- sal_long %>%
  filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  filter(district == "sesi") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(SpCode, Mon, Habitat) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity") +
  plot_theme() 
sesi1 + facet_wrap(~SpCode, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```


# Next SESO (Southeast Southern Outside - eastern POW/Exposed Ketchikan/Gravina Island area)

```{r seso.hab, echo = FALSE, fig.width=10, fig.height=6, fig.cap=fig$cap("seso.hab", "Number of salmon caught in summer months at four different habitat types in Southeast Southern Outside area with free y axis scale"), message=FALSE, warning=FALSE}
sesi1 <- sal_long %>%
  filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  filter(district == "seso") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(SpCode, Mon, Habitat) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity") +
  plot_theme() 
sesi1 + facet_wrap(~SpCode, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```


# Combine pink salmon graphs

Lets look at only pink salmon for a second across each district

```{r pink, echo = FALSE, fig.width=8, fig.height=6, fig.cap=fig$cap("pink", "Number of pink salmon caught in summer months at four different habitat types with free y axis scales across four Southeastern districts excluding Yakutat"), message=FALSE, warning=FALSE}
pink <- sal_long %>%
  filter(SpCode == "SALPINK") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Habitat, district) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity", position = "dodge") +
  plot_theme() + ylab("Total Pink Salmon") + xlab("Months")
pink + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

You'll notice that the patterns of more pink salmon in bedrock we don't necessarily see even in the SESO district (that we based the glm models on). The reason for this is that for SESO sites, I included **only sites I could pair**. Meaning I restricted the analysis to sites that were seined at the same time and that had vegetated and unvegetated habitats directly next to or closely adjacent. 

# Chum salmon
```{r chum, echo=FALSE, fig.cap=fig$cap("chum", "Number of chum salmon caught in summer months at four different habitat types with free y axis scales across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
chum <- sal_long %>%
  filter(SpCode == "SALCHUM") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Habitat, district) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity", position = "dodge") +
  plot_theme() + ylab("Total Chum Salmon") + xlab("Months")
chum + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```


# Coho salmon
```{r coho, echo=FALSE, fig.cap=fig$cap("coho", "Number of coho salmon caught in summer months at four different habitat types with free y axis scales across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=8, message = FALSE, warning=FALSE}
coho <- sal_long %>%
  filter(SpCode == "SALCOHO") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Habitat, district) %>%
  dplyr::summarise(total = sum(abundance)) %>%
  ggplot() + geom_bar(aes(x = Mon, y = total, fill = Habitat), stat = "identity", position = "dodge") +
  plot_theme() + ylab("Total Coho Salmon") + xlab("Months")
coho + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

Okay so clearly there's just a ton of variability in total salmon abundances. Lets look in average numbers because there definitely were habitats that were seined more than others. 

# Average salmon caught by hab 
## Pink salmon
```{r avg.pink, echo=FALSE, fig.cap=fig$cap("avg.pink", "Average (+sd) number of pink salmon caught in summer months at four different habitat types with free y axis scales across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=8, warning=FALSE, message = FALSE}
avg.pink <- sal_long %>%
  filter(SpCode == "SALPINK") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Habitat, district) %>%
  dplyr::summarise(avg = mean(abundance), sd = sd(abundance)) %>%
  ggplot(aes(x = Mon, y = avg, ymin = avg, ymax = avg+sd, fill = Habitat)) + 
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(position = position_dodge(width = 0.9), colour = "black", width = 0.1) +
  plot_theme() + ylab("Average Pink Salmon") + xlab("Months")
avg.pink + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

Only included the positive standard deviation in `r fig$ref("avg.pink")` because it got real messy/ugly with the negative standard deviation. Basically, by visual assesment there doesn't look like true difference between the habitats. Remember y axis is free. 

## Chum salmon
```{r avg.chum, echo=FALSE, fig.cap=fig$cap("avg.chum", "Average (+sd) number of chum salmon caught in summer months at four different habitat types with free y axis scales across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=8, warning=FALSE}
avg.chum <- sal_long %>%
  filter(SpCode == "SALCHUM") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Habitat, district) %>%
  dplyr::summarise(avg = mean(abundance), sd = sd(abundance)) %>%
  ggplot(aes(x = Mon, y = avg, ymin = avg, ymax = avg+sd, fill = Habitat)) + 
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(position = position_dodge(width = 0.9), colour = "black", width = 0.1) +
  plot_theme() + ylab("Average Chum Salmon") + xlab("Months")
avg.chum + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(name = "Habitat", palette = "Dark2")
```

# Veg/Unveg
What if we looked strictly at vegetated versus unvegetated
## Pink salmon
```{r veg.unveg, echo=FALSE, fig.cap=fig$cap("veg.unveg", "Average (+sd) number of pink salmon caught in summer months in vegetated and unvegetated habitats with free y axis scales across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=8, warning=FALSE, message = FALSE}

pink2 <- sal_long %>%
  filter(SpCode == "SALPINK") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  mutate(Mon = fct_relevel(Mon, 
                           c("Apr", "May", "Jun", "Jul"))) %>%
  group_by(Mon, Hab_type, district) %>%
  dplyr::summarise(avg = mean(abundance), sd = sd(abundance)) %>%
  ggplot(aes(x = Mon, y = avg, ymin = avg, ymax = avg+sd, fill = Hab_type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(position = position_dodge(width = 0.9), colour = "black", width = 0.1) +
  plot_theme() + ylab("Average Pink Salmon") + xlab("Months")
pink2 + facet_wrap(~district, scales = "free_y") + scale_fill_brewer(palette = "Dark2")

```

### Boxplot version
Lets look at the boxplot version of that
```{r boxvu, echo=FALSE, fig.cap=fig$cap("boxvu", "Boxplot of pink salmon abundances across four Southeastern districts excluding Yakutat"), fig.height=6, fig.width=7, warning=FALSE}

boxvu <- sal_long %>%
  filter(SpCode == "SALPINK") %>%
    filter(Mon == "Apr" |
           Mon == "May" |
           Mon == "Jun" |
           Mon == "Jul") %>%
  group_by(Hab_type, district) %>%
  ggplot(aes(x = Hab_type, y = abundance)) + 
  geom_boxplot() +
  plot_theme() + ylab("Pink Salmon Abundance") + xlab("Habitat")
boxvu + facet_wrap(~district, scales = "free_y")

```

Theres so much variability! This graph only includes those same four summer months again. 



