---
title: "ShoreZone mapping"
author: "Lia Domke"
date: "10/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Interested in trying to use shorezone in R
Want to:

+ input shapefiles clipped to seak
+ subset to habitat types of interest


The final goal is looking at the number of salmon that occur in primarily two "comparitive" vegetated habitats. 
ShoreZone using aerial imagery classified major habitat types, including Zostera marina (Zos) and soft brown kelps (SBR). However, based on the biological classification and habitat and exposure classification it was difficult/impossible to quanitfy just bedrock/sand-gravel habitats. 
So I want to determine if salmon preferentially occur in eelgrass over soft brown kelps. 
Needs:
- total number of salmon caught in eelgrass and understory kelp habitats
- proportion caught in each habitat
- km of eelgrass within the "SESO" region 
- km of soft brown kelps within the "SESO" region


Will need diff than normal libraries for mapping
```{r}
library(rgdal)
library(ggplot2)
library(rgeos)
library(raster)
library(maptools)
library(sf)
library(dplyr)
library(tidyr)
```

Input ShoreZone layers
  Notes about shorezone layers: Through the new shorezone java site (https://alaskafisheries.noaa.gov/mapping/sz_js/index.html?tab=sz&layout=h2#) I downloaded the whole shorezone .mpk file that included shorezone data for Alaska and part of oregon. I changed the extension to .zip and unzipped it in order to load it into qgis (could also use arcgis, but didn't have access-qgis is an open source mac compatible alterative to arcgis). 
  In qgis (version 3.10.10) I loaded the AK_UNIT_W_ATTRS_W_XSHR shapefile within the v10 file folder. This shapefile had attributes for major biological, habitat, and exposure attributes among others. For attribute definitions and additional information consult shorezone metadata found on the shorezone java site. The file reads in with an inbedded CRS (espg3857), documentation indicates that it is in WGS 1984 Major Auxiliary Sphere. 
  I then created a polygon in qgis (SESO clipper_1-7-21_espg3857) to clip the large shapefile (extent all of Alaska) to just approximately the expanse of the Southeast Southern Outside waters as defined by the NOAA Nearshore Atlas data. All further analysis in this script will be based on that area. 
  I then projected that clipped lines data into a project CRS appropriate (I think) for this area. I used NAD83(2011) Alaska zone 1 espg 6394 for the SE Panhandle. (file name AK_SESO_espg6394).
  I used vector geometry toolbox to add geom attributes (making sure to use ellopsial rather than the layer CRS), this added length to the different attributes ("length" in meters). I then was able to calculate the total coastline, continuous zostera coastline, and the continuous soft (non-stalked) brown kelp coastline. 
```{r}
# read in a feature class
seak_base <- readOGR(dsn = "../Maps_NOAA&Otter",
        layer = "southern_seak_espg6394")

attr <- readOGR(dsn = "../Maps_NOAA&Otter", 
                   layer = "AK_LINES_SESO_espg6394_geom")
par(mfrow=c(1,1))
```

In order to filter shapefiles with dplyr have to transform to SF object and then filter
```{r}
attr_sf <- st_as_sf(attr)

# filter data by zostera/soft brown kelp only
zos <- attr_sf %>%
  filter(ZOS == "C")

sbr <- attr_sf %>%
  filter(SBR == "C")

# lets look at this mapped real quick, transform back to Spatial class
zos_sp <- as(zos, Class = "Spatial")
sbr_sp <- as(sbr, Class = "Spatial")
attr_sp <- as(attr_sf, Class = "Spatial")
```

Want to measure total distance of the zostera spatiallinesdataframe
```{r}
summary(zos_sp)

st_length(zos_sp)
zos_km <- sum(st_length(zos))/1000 # lets use this one

# lets transform the data an dsee how its different
zos_utm <- spTransform(zos_sp, CRS("+init=epsg:4326"))
zos_utm_sf <- st_as_sf(zos_utm)
sum(st_length(zos_utm_sf))/1000 # in kilometers! #

# Zos_km is the same exact measurement that I got with the qGIS geom attributes based analysis (w/ espg 6394)! But not the same as the intergrated shorezone measurement. The shorezone measurement was 840.08 km (difference of 140 km). However if I use what proportion of the caostline is eelgrass or understory kelp and make sure I use the same method of measurement it should be okay. 

# Soft brown kelp coastline
st_length(sbr_sp)
sbr_km <- sum(st_length(sbr))/1000 # lets use this one

# total coastline
st_length(attr_sp)
coast_km <- sum(st_length(attr_sf))/1000
```

Lets plot the diff maps real quick - this takes a long time, don't do if we can avoid.
```{r}
# high light all three lines at the same time. 
#plot(seak_base, col = "grey")
#plot(sbr_sp,  col = "brown", add = TRUE)
#plot(zos_sp, col = "green", add = TRUE)


# plot(attr_sp)
```

Proportion of each coastline type
```{r}
zos_perc <- zos_km/coast_km* 100
sbr_perc <- sbr_km/coast_km* 100
```

Initially lets just use data from the Nearshore Atlas seines
We want to make sure we're only including data from the SESO region and that we look at seines that occurred between April and July (based on when salmon outmigrated in 2017). We also only want seines that happened in eelgrass and understory kelp habitats. 
```{r data}
seso <- read.csv("Data/noaa_region_seso.csv", stringsAsFactors=FALSE, header = TRUE)
```

filter by hab
```{r}
glimpse(seso)

seso_fish <- seso %>%
  filter(Habitat == "Eelgrass"| Habitat == "Kelp") %>%
  filter(taxon == "fish")
```

filter by species (salmon only)
```{r}
seso_sal <- seso_fish %>%
  filter(Gear == "BSEINE") %>%
  filter(SpCode == "SALSOCK"| SpCode == "SALCHIN" | SpCode == "SALPINK"| 
           SpCode == "SALCOHO" | SpCode == "SALCHUM")
```

Graph species occurrence by month
```{r}
seso_sal$abundance <- as.numeric(ifelse(is.na(seso_sal$Length), paste(seso_sal$Unmeasured), 1))

seso_sal %>%
  ggplot() +
  geom_bar(mapping = aes(y = abundance, x = Mon, fill = SpCode), stat = "identity")
```

There were fish caught in April, May, June, July (latest seine in July is 7/16/1999) so for now lets keep those four months
filter by month
```{r}
seso_sal <- seso_sal %>%
  filter(Mon == "Apr"|Mon == "May"|Mon == "Jul"|Mon=="Jun")
```

What was the total number of salmon caught during this time frame in these two habitats
```{r}
sum <- seso_sal %>%
  group_by(SiteID, EventID, Date, Mon, Habitat, SpCode) %>%
  dplyr::summarise(total_abund = sum(abundance))
```

Based on raw numbers: 
```{r}
x <- sum %>%
  ungroup() %>%
  dplyr::summarise(sum = sum(total_abund)) # 11938 total number of salmon caught

total_sal <- x$sum

x1 <- sum %>%
  ungroup() %>%
  filter(Habitat == "Eelgrass") %>%
  dplyr::summarise(sum = sum(total_abund))

x2 <- sum %>%
  ungroup() %>%
  filter(Habitat == "Kelp") %>%
  dplyr::summarise(sum = sum(total_abund))

zos_sal <- x1$sum # number of salmon caught in zostera
sbr_sal <- x2$sum # number of salmon caught in understory kelp

# number of salmon expected to be caught based on the percent of coastline each habitat occupies
exp_zos <- as.vector(total_sal * (zos_perc/100)) # as.vector removes the units
exp_sbr <- as.vector(total_sal * (sbr_perc/100))
```
Lets also calculate based on average numbers to account for sampling effort in each hab
```{r}
# want number of seines in each hab during this time frame
glimpse(sum)

n_in_eel <- sum %>%
  filter(Habitat == "Eelgrass") %>%
  ungroup() %>%
  dplyr::select(EventID, Date) %>%
  distinct() %>%
  dplyr::summarise(n()) # total number of salmon caught in seines (all seines all species May - July) in eelgrass 

n_in_sbr <- sum %>%
  filter(Habitat == "Kelp") %>%
  ungroup() %>%
  dplyr::select(EventID, Date) %>%
  distinct() %>%
  dplyr::summarise(n()) # total number of salmon caught in seines (all seines all sepcies, May - July) in understory kelp

avg_eel_sal <- sum %>%
  filter(Habitat == "Eelgrass") %>%
  group_by(EventID, Date) %>%
  dplyr::summarise(sum_by_eventID = sum(total_abund)) %>%
  ungroup() %>%
  dplyr::summarise(mean(sum_by_eventID)) # average number of salmon per seine in eeelgrass hab #~ 359.46

avg_sbr_sal <- sum %>%
  filter(Habitat == "Kelp") %>%
  group_by(EventID, Date) %>%
  dplyr::summarise(sum_by_eventID = sum(total_abund)) %>%
  ungroup() %>%
  dplyr::summarise(mean(sum_by_eventID)) # average number of salmon per seine in kelp hab# ~81.43 

avg_sal <- sum %>%
  group_by(EventID, Date) %>%
  dplyr::summarise(sum_by_eventID = sum(total_abund)) %>%
  ungroup() %>%
  dplyr::summarise(mean(sum_by_eventID))
```

Lets step up a dataframe that has all this info
```{r}
habitat <- c("eelgrass", "kelp")
expected_perc <- c(as.vector(zos_perc), as.vector(sbr_perc))
expected_freq <- c(exp_zos, exp_sbr)
observed_freq <- c(zos_sal, sbr_sal)
n <- c(n_in_eel$`n()`, n_in_sbr$`n()`)

dat <- data.frame(habitat, expected_perc, expected_freq, observed_freq, n)
```

One thing that isn't accounted for in this is the sampling effort. 
If we know that 11938 salmon were caught in 51 seines in two habitats, then based on our 
H0: if the salmon occur proportionally in eelgrass and kelp habitats then the average number of salmon would be equal to --> 
total_salmon(prop_of_hab)/# of seines that happened in that habitat

H1: if the salmon do not occur proportionally the average number of salmon would not be equal the 
eq above

But is this complicating this too much?

Eelgrass composes 14% of the coastline that was sampled and soft brown kelps compose 39% of the coastline. The two habitats *can* overlap so there could be over 100% of habitat 'sampled'.
28 eelgrass sites and 23 kelp sites were sampled between April and July in 1998 and 1999. 

First lets look at average salmon catch in each habitat type in seso
```{r theme settings, include=FALSE}
# Creates custom base plot theme that can adjust every graph that you use plot_theme for!

plot_theme <- function() {
  theme_bw(base_size = 48, base_family = "Avenir Black") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="transparent", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank())
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```
```{r}
sum %>%
  group_by(Habitat) %>%
  dplyr::summarise(avg = mean(total_abund), sd = sd(total_abund)) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x = Habitat, y = avg), stat="identity") + 
  geom_errorbar(mapping = aes(x = Habitat, ymax = avg+sd, ymin = avg-sd), 
                width = 0, stat = "identity") +
  plot_theme()
```

Lets explore the data briefly, is it normally distributed, does it have equal variances?
```{r}
hist(sum$total_abund)
hist((sum$total_abund)^(1/4))
hist(sqrt(sum$total_abund))
hist(log(sum$total_abund))

boxplot(sum$total_abund ~ sum$Habitat)
```
Nah they really don't


We know that we are comparing the means of categorical variables (indep) against continuous variables (dept) and that the data is nonparametric (i.e. not normal). Should compare means using Mann-Whitney test

Four assumptions of the test:
1. categorical is independent
2. dependent is continuous
3. independence of observations (no correlation between samples)
4. Need to determine distributins of both categories (in this case hab) 
have the same shape or different shape

```{r}
sum %>%
  ggplot() +
  geom_density(aes(x= (total_abund), color = Habitat)) +
  plot_theme()
# hard to see the distribution and if they're similar or not 

# log distribution
sum %>%
  ggplot() +
  geom_density(aes(x= log(total_abund), color = Habitat)) +
  plot_theme()
```
The distributions of the data by group appear to be similar. They both seem to be positively skewed both if they are log transformed or not. In this case the Mann-Whitney test, *the median can be used to assess the differences in Habitats*

H0: There is no difference between salmon abundances for eelgrass compared to understory kelp 
```{r}
# prep the data take out species covariate
dat2 <- sum %>%
  group_by(EventID, Habitat) %>%
  dplyr::summarise(abund_by_eventID = sum(total_abund)) # 51 total events
dat3 <- sum %>%
  filter(Mon != "Jul") %>% # filter out july seines
  group_by(EventID, Habitat) %>%
  dplyr::summarise(abund_by_eventID = sum(total_abund)) 
  
dat2$Habitat<-as.factor(dat2$Habitat)

wilcox.test(abund_by_eventID ~ Habitat, data = dat2) # defaults to mannU test
wilcox.test(abund_by_eventID ~ Habitat, data = dat3) # data w/o july seines
```

so based on the Mann-Whitney test there is no difference in the median of salmon abundance between eelgrass and understory kelp. 

Okay so our question then is:
* When salmon are present, do they occur more in eelgrass or understory kelp? * 
H0: structured habitats regardless of type are equally important for salmon
H1: one structured habitat has more salmon occur in the habitat. So we know that understory kelp occurrs in 39% of the coastline studied and zostera in 14% but thats not equal to 100% so there is some overlap in sbr and zos line features. 

In order to look at exclusively one habitat areas (just eelgrass OR softbrown kelps):
I went back to QGIS and 'selected features by expression' and did a two part selection. First I selected features where ZOS == 'C' giving selected features of all the continuous zostera. Second I removed all features that were SBR == 'C' resulting in just zostera where soft brown kelps do NOT occur in a continuous way. 

And then I did the same thing for soft brown kelps (sbr)
```{r}
zos_wo_sbr <- 761.489 # km of zostera that aren't with sbr
sbr_wo_zos <- 2556.18 # km of sbr that aren't overlapping with zos

ttl <- zos_wo_sbr + sbr_wo_zos # total coastline of either EELGRASS OR SOFTBROWN KELPS

zos_wo_sbr_perc <- zos_wo_sbr/ttl*100
sbr_wo_zos_perc <- sbr_wo_zos/ttl*100

zos_wo_sbr/coast_km
sbr_wo_zos/coast_km
ttl/coast_km
```

Because we have salmon numbers from *either* eelgrass or understory kelp, lets base our calculations on just the area that has either habitat (not both)

In terms of numbers, this analysis of habitats covers 47% of the total coastline in southern southeast alaska. Just Zostera is 10.8% of the *total* coastline, sbr is 36.5% of the *total* coastline. 
```{r}
habitat <- c("eelgrass", "kelp")
expected_perc2 <- c(as.vector(zos_wo_sbr_perc/100), as.vector(sbr_wo_zos_perc/100))
expected_freq2 <- c((x$sum * (zos_wo_sbr_perc/100)), (x$sum *(sbr_wo_zos_perc/100)))
observed_freq2 <- c(zos_sal, sbr_sal)
n2 <- c(n_in_eel$`n()`, n_in_sbr$`n()`)

dat_singhab <- data.frame(habitat, expected_perc2, expected_freq2, observed_freq2, n2)
```

```{r}
dat_singhab %>%
  dplyr::select(-n2,-expected_perc2) %>%
  pivot_longer(cols=expected_freq2:observed_freq2) %>%
  ggplot() +
  geom_bar(aes(x = name, y = value, fill = habitat), stat = "identity") +
  ylab("Number of Salmon Caught") + xlab(NULL) +
  scale_x_discrete(labels=c("expected_freq2" = "Expected Frequency", 
                            "observed_freq2" = "Observed Frequency")) +
  scale_fill_manual(values = c("#339900", "#D16103"), name = "Habitat", 
                   labels = c("Eelgrass", "Understory Kelp")) +
  plot_theme()
```

Set up the chi square test
```{r}
chisq.test(dat_singhab$observed_freq2,p=dat_singhab$expected_perc2)
```

