---
title: "Chum & Pink SAV use analysis objective 1"
author: "Lia Domke"
date: "9/1/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

Juvenile Chum and Pink Salmon use of submerged vegetative habitats in Southeast Alaska in Marine and Coastal Fisheries 

Contributed by Lia K Domke, Rebecca J Cates, Wendel W Raymond, Ginny L Eckert

Specifically analyses related to **objective 1** in the manuscript

1) Investigate chum (O. keta) and pink (O. gorbuscha) abundance in nearshore marine habitats (eelgrass and understory kelp)

# Cleaning salmon data
Data from the Nearshore Atlas seines

We want to make sure we're only including data from the SESO region and that we look at seines that occurred between April and July (based on when salmon outmigrated in 2017). We also only want seines that happened in eelgrass and understory kelp habitats. 

```{r data, include=FALSE}
seso <- read.csv("Data/noaa_region_seso.csv", stringsAsFactors=FALSE, header = TRUE)
```

#Libraries
```{r}
library(tidyverse)
library(glmmTMB)
library(AICcmodavg)
library(pscl) # vuong test
library(visreg)
#library(ggsignif) 
library(patchwork)
library(mgcv)
```

1. Fix incorrect species

```{r include=FALSE}
unique(seso$taxon)
unique(seso$Sp_CommonName) # there are a couple of non fish species, 93 unique species 

incor <- c("SHRMGRA", "BSNAILW", "CRABHEL", "JLYMOON", "JLYCLNG", "TUNICAT", "SHRMDOC", "STARLEA", "SHRMSTI",
           "SHRMSPO", "CRABHER", "SHRMHIP", "ISOPODG", "CRABGD", "CRABDUN", "SHRMCR", "ISOPODC", "SHRMCOO",
           "CRABGKE", "UNMYSID", "CRABNKE", "NUDOPAL", "NUDHOOD")

seso_fish <- seso %>%
  mutate(taxon = ifelse(SpCode %in% incor, "invertebrate", taxon)) %>%
  filter(taxon == "fish") %>%
    filter(Habitat == "Eelgrass"| Habitat == "Kelp")

#unique(seso_fish$Sp_CommonName) # now only 70 
```

2. Filter by habitat

```{r include=FALSE}
glimpse(seso_fish)

seso_fish_sub <- seso_fish %>%
  filter(Mon == "Jul" | Mon == "Apr" | Mon == "Jun" | Mon == "May") %>%
  filter(Gear == "BSEINE")

sites_by_event <- unique(seso_fish_sub[c("SiteID", "EventID","Habitat", "Date", "Mon", "Year", "Locale", "Location")])
length(unique(seso_fish_sub$SiteID)) # 42 sites that were in southern southeast alaska in the Months of interest
```

Create sal.wide.sp/sal.long.sp that INCLUDES species - this should be used for species specific modelling.
```{r}
sal.wide.sp <- seso_fish_sub %>%
  mutate(abundance = as.numeric(ifelse(is.na(seso_fish_sub$Length), paste(seso_fish_sub$Unmeasured), 1))) %>%
  pivot_wider(id_cols = c("SiteID", "EventID","Habitat", "Date", "Mon", "Year", "Locale", "Location",
                          "Lat1", "Long1"), names_from=SpCode, values_from=abundance, values_fill = 0, values_fn = sum) %>%
  dplyr::select(c("SiteID", "EventID","Habitat", "Date", "Mon", "Year", "Locale", "Location",
                          "Lat1", "Long1", "SALSOCK", "SALPINK", "SALCOHO", "SALCHUM")) # No chinook in the data

sal.long.sp <- sal.wide.sp %>%
  pivot_longer(cols = c("SALSOCK", "SALPINK", "SALCOHO", "SALCHUM"), names_to = "SpCode", values_to = "abundance")

#write.csv(sal.long.sp, "chp1.sal.long.sp_9-1-23.csv")
```

# Custom fxns
## Plot theme
```{r theme settings, include=FALSE}
# Creates custom base plot theme that can adjust every graph that you use plot_theme for! First create custom plot theme

plot_theme <- function() {
  theme_bw(base_size = 24, base_family = "Helvetica") %+replace%
    theme(panel.background  = element_blank(),
            plot.background = element_rect(fill="transparent", colour=NA), 
            legend.background = element_rect(fill="transparent", colour=NA),
            legend.key = element_rect(fill="transparent", colour=NA),
            panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank())
}

# to use ggsave and export plots include argument 'device=cario_pdf' e.g.: 
# ggsave("name.pdf", device=cairo_pdf, width = 6, height = 6)
```

## global label fxn
```{r adding global label}
add_global_label <- function(pwobj, Xlab = NULL, Ylab = NULL, Xgap = 0.03, Ygap = 0.03, ...) {
    ylabgrob <- patchwork::plot_spacer()
    if (!is.null(Ylab)) {
        ylabgrob <- ggplot() +
            geom_text(aes(x = .5, y = .5), label = Ylab, parse = T,
                      angle = 90, ...) +
            theme_void()
    }
    if (!is.null(Xlab)) {
        xlabgrob <- ggplot() +
            geom_text(aes(x = .5, y = .5), label = Xlab, ...) +
            theme_void()
    }
    if (!is.null(Ylab) & is.null(Xlab)) {
        return((ylabgrob + patchworkGrob(pwobj)) + 
            patchwork::plot_layout(widths = 100 * c(Ygap, 1 - Ygap)))
    }
    if (is.null(Ylab) & !is.null(Xlab)) {
        return((ylabgrob + pwobj) + 
            (xlabgrob) +
            patchwork::plot_layout(heights = 100 * c(1 - Xgap, Xgap),
                                   widths = c(0, 100),
                                   design = "
                                   AB
                                   CC
                                   "
            ))
    }
    if (!is.null(Ylab) & !is.null(Xlab)) {
        return((ylabgrob + pwobj) + 
            (xlabgrob) +
            patchwork::plot_layout(heights = 100 * c(1 - Xgap, Xgap),
                                   widths = 100 * c(Ygap, 1 - Ygap),
                                   design = "
                                   AB
                                   CC
                                   "
            ))
    }
    return(pwobj)
}

```


# Objective 1
Also - you need to create box plots of salmon catch between month and eelgrass type


We're comparing salmon counts by species (Chum and Pink) across habitat types and incorporating sampling timeframe as well as random effects of SiteID because we have 12 sites that we're resampled

Following Diggle et al 2002 "top-down" strategy
1. test random effects structure (this hasto be done using method = "REML")

Sal ~ Habitat + (1|Mon) + (1|SiteID)
Sal ~ Habitat + (Mon|SiteID)
Sal ~ Habitat + Mon + (1|SiteID)

Compare the model and random structure using AIC, BIC, likelihood ratio test

2. Switch to method = "ML", and determine fixed effects
We only have two potential fixed effects (month and/or habitat)
And we're truly interested in Habitat and Month is important thing we want to account for. In the end it'll include both these factors

3. Refit ideal random and fixed effect structure with method = 'REML'

We're going to use glmmTMB and see if we can get generalized mixed models for both the negative binomial and zeroinflated NB to work

# Seperate species
```{r}
glimpse(sal.long.sp)

chum.long <- sal.long.sp %>%
  filter(SpCode == "SALCHUM")

pink.long <- sal.long.sp %>%
  filter(SpCode == "SALPINK")
```
Also - you need to create box plots of salmon catch between month and eelgrass type

# Sp boxplots by month (Supp figure 3)
```{r}
p1 <- sal.long.sp %>%
   mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  filter(SpCode == "SALCHUM") %>%
  ggplot() +
  geom_boxplot(aes(x = Mon, y = abundance, fill = Habitat)) +
  plot_theme()  +
  scale_fill_manual(values = c("#327610", "#F46036"),
                    labels = c("Eelgrass", "Understory kelp"))+
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  labs(x = "Month", y = Number~of~salmon~(no.~seine^-1)) +
  theme(legend.position = "none")

p1

p2 <- sal.long.sp %>%
   mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  filter(SpCode == "SALPINK") %>%
  ggplot() +
  geom_boxplot(aes(x = Mon, y = abundance, fill = Habitat)) +
  plot_theme()  +
  scale_fill_manual(values = c("#327610", "#F46036"),
                    labels = c("Eelgrass", "Understory kelp"))+
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  labs(x = "Month", y = NULL) +
  theme(legend.position = "none")

p2

plot <- p1 + p2 + plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect") & theme(legend.position = "bottom", 
                                          legend.key.size = unit(1.5, "cm"))
plot

#ggsave("../../Chp1_salmon/Manuscript_scripts&data/chp1_obj1_nfa_abund_plots091724.png", plot = plot, width = 14, height = 9)
```

# Random effects structure testing both species
## Chum RE sturcture testing
```{r}
ch.re1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re1)

ch.re2 <- glmmTMB(abundance ~ Habitat + (1|Mon) + (1|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re2)

ch.re3 <- glmmTMB(abundance ~ Habitat + (Mon|SiteID), data = chum.long, REML = T, family = nbinom2(link = "log"))
summary(ch.re3)

AICc(ch.re1); AICc(ch.re2); AICc(ch.re3)
BIC(ch.re1); BIC(ch.re2); BIC(ch.re3)
# the best random effects model is the mon as a fixed factor
```

## Pink RE structure testing
```{r}
p.re1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re1)

p.re2 <- glmmTMB(abundance ~ Habitat + (1|Mon) + (1|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re2)

p.re3 <- glmmTMB(abundance ~ Habitat + (Mon|SiteID), data = pink.long, REML = T, family = nbinom2(link = "log"))
summary(p.re3)

AICc(p.re1); AICc(p.re2); AICc(p.re3)
BIC(p.re1); BIC(p.re2); BIC(p.re3)
```

Okay for the two species - having a random effects structure of slope rather than just random intercept structure is not possilbe 

For both species the random intercept of site is the best form using minimized AIC and BIC

2. Switch to method = "ML", and determine fixed effects
We only have two potential fixed effects (month and/or habitat)
And we're truly interested in Habitat and Month is important thing we want to account for
In the end it'll include both these factors


# Chum fixed effects structure
## Chum glmmTMB
```{r}
chum1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long,
                 REML = F, family = nbinom2(link = "log"))
summary(chum1)

chum2 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = chum.long[-26,],
                 REML = F, family = nbinom2(link = "log"))
summary(chum2)

# all these are significant so this is the best model? especially since I decided that doy or month wasn't actually continuous rather clustered within certain time frames
# woo here's the chum model
AICc(chum1); AICc(chum2)
BIC(chum1); AICc(chum2)
# we should test the zero inflated model tho too

par(mfrow=c(2,2))
library(DHARMa)
plot(simulateResiduals(chum1))
plot(simulateResiduals(chum2))
```

## Chum gam w/ space
```{r}
chum1.gam <- gam(abundance ~ Habitat + Mon + te(Lat1, Long1, k = 3) + 
                   s(SiteID, bs = "re"), 
                 data = chum.long,
                 REML = F, family = nb())
summary(chum1.gam)

chum2.gam <- gam(abundance ~ Habitat + Mon + s(Lat1, Long1) + 
                   s(SiteID, bs = "re"), 
                 data = chum.long,
                 REML = F, family = nb())
summary(chum2.gam)

chum3.gam <- gam(abundance ~ Habitat + Mon + s(Lat1) + 
                   s(SiteID, bs = "re"), 
                 data = chum.long,
                 REML = F, family = nb())
summary(chum3.gam)

chum4.gam <- gam(abundance ~ Habitat + Mon + s(Lat1), 
                 data = chum.long,
                 REML = F, family = nb())
summary(chum4.gam)

chum5.gam <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
                 data = chum.long,
                 REML = F, family = nb())
summary(chum5.gam)

AICc(chum1.gam); AICc(chum2.gam); AICc(chum3.gam); AICc(chum4.gam); AICc(chum5.gam)
# best model is chum4.gam which includes habitat month and smoothed lat
```

## is zero inflated better?
Okay following the vignette about glmmTMB we can fit zero inflated negative binomial models using ziformula = ~1

We cant compare with the s(lat) as far as I can tell because we can't include smoothed terms in glmmTMB and we can't use the zero inflated or hurdle model with gams

```{r}
chum.zinf <- glmmTMB(abundance ~ Habitat + Mon + Lat1, data = chum.long, 
                 REML = F, family = nbinom2(link = "log"),
                 ziformula = ~1)
summary(chum.zinf)

chum.hurd <- glmmTMB(abundance ~ Habitat + Mon + Lat1, data = chum.long, 
                 REML = F, family = truncated_nbinom2(link = "log"),
                 ziformula = ~1)
summary(chum.hurd)

# what if we included tweedie
chum3.gam.tw <- gam(abundance ~ Habitat + Mon + s(Lat1) + s(SiteID, bs = "re"), 
                    data = chum.long,
                    family = tw(link = "log"))
chum3.glm.tw <- glm(abundance ~ Habitat + Mon + Lat1 + (1|SiteID), data = chum.long,
                       family = statmod::tweedie(link.power = 0, var.power =1.632))


AICc(chum3.gam); AICc(chum.zinf); AICc(chum.hurd); AICc(chum3.gam.tw)
BIC(chum3.gam); BIC(chum.zinf); BIC(chum.hurd)

tweedie::AICtweedie(chum3.glm.tw); AIC(chum3.gam); AIC(chum.zinf); AIC(chum.hurd); AIC(chum3.gam.tw) # best model is still the neg bin gam

AICc(chum.zinf) - AICc(chum3.gam)
AICc(chum3.gam.tw) - AICc(chum3.gam)

# vuong test compares two non-nested hypothesis test
# good way of comparing model formulas that are non nested such as zero inflated and non-zero inflated analogs
# vuong(chum4.gam, chum.zinf)
# okay I dont think this works with random effects models, so we'll rely on AIC/BIC values

performance::check_overdispersion(chum3.gam)

plot(simulateResiduals(chum3.gam.tw))
plot(simulateResiduals(chum3.gam))
```

## testing outliers w/ best model
```{r}
chum.long.red1 <- chum.long %>%
  filter(abundance < 100) # removing all occurence greater than 100, n = 71 obs

chum4.gam.red1 <- gam(abundance ~ Habitat + Mon + s(Lat1), 
                 data = chum.long.red1,
                 REML = F, family = nb())
summary(chum4.gam.red1)

plot(simulateResiduals(chum4.gam))
plot(simulateResiduals(chum4.gam.red1))
```


# Pink Fixed effects

## Pink gam (updated approach 9/24)
```{r}
pink1.gam <- gam(abundance ~ Habitat + Mon + te(Lat1, Long1, k = 3) + 
                   s(SiteID, bs = "re"), 
                 data = pink.long,
                 REML = F, family = nb())
summary(pink1.gam)

pink2.gam <- gam(abundance ~ Habitat + Mon + s(Lat1, Long1) + 
                   s(SiteID, bs = "re"), 
                 data = pink.long,
                 REML = F, family = nb())
summary(pink2.gam)

pink3.gam <- gam(abundance ~ Habitat + Mon + s(Lat1) + 
                   s(SiteID, bs = "re"), 
                 data = pink.long,
                 REML = F, family = nb())
summary(pink3.gam)

pink4.gam <- gam(abundance ~ Habitat + Mon + s(Lat1), 
                 data = pink.long,
                 REML = F, family = nb())
summary(pink4.gam)

pink5.gam <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
                 data = pink.long,
                 REML = F, family = nb())
summary(pink5.gam)

AICc(pink1.gam); AICc(pink2.gam); AICc(pink3.gam); AICc(pink4.gam); AICc(pink5.gam)
# best model is chum5.gam which includes includes habitat month and re site
```

##Pink glmmTMB
```{r}
pink1 <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, REML = F, family = nbinom2(link = "log"))
summary(pink1)

pink2 <- glmmTMB(abundance ~ Habitat + Mon, data = pink.long, REML = F, family = nbinom2(link = "log"))
summary(pink2)

pink.long.red <- pink.long[-c(22,26,46,19,40),]
pink1.red <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), 
                     data = pink.long.red, REML = F, 
                     family = nbinom2(link = "log"))

AICc(pink1); AICc(pink2); AICc(pink1.red)
BIC(pink1); BIC(pink2)
# the one without random effects is slightly better; however, we know there is structure in it so i'm not inclined to drop it.

plot(simulateResiduals(pink1))
plot(simulateResiduals(pink1.red))
```

## testing zero inflated
```{r}
pink.zinf <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, 
                 REML = F, family = nbinom2(link = "log"),
                 ziformula = ~1)
summary(pink.zinf)

pink.hurd <- glmmTMB(abundance ~ Habitat + Mon + (1|SiteID), data = pink.long, 
                 REML = F, family = truncated_nbinom2(link = "log"),
                 ziformula = ~1)
summary(pink.hurd)

pink5.gam.tw <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
                 data = pink.long,
                 REML = F, family = tw())
summary(pink5.gam.tw)

performance::check_overdispersion(pink5.gam)
# can't actually fit the tweedie with glm because of the re effects
#pink5.glm.tw <- glm(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
#                 data = pink.long,
#                 REML = F, family = statmod::tweedie(var.power = 1.759, link.power = 0))
#summary(pink5.glm.tw)

AICc(pink4.gam); AICc(pink.zinf); AICc(pink.hurd); AICc(pink5.gam.tw)
BIC(pink4.gam); BIC(pink.zinf); BIC(pink.hurd); BIC(pink5.gam.tw)
AIC(pink4.gam); AIC(pink.zinf); AIC(pink.hurd); AIC(pink5.gam.tw)
# best model is the gam form of habitat + mon + site re

plot(simulateResiduals(pink4.gam))
plot(simulateResiduals(pink5.gam.tw))

AICc(pink5.gam.tw) - AICc(pink4.gam)
```


For both species the normal negative binomial reduces AICc / BIC values more than the zero inflated negbin or the hurdle model 

3. Refit ideal random and fixed effect structure with method = 'REML'

## testing outliers w/ best model
This means we have to remove abundance values above like 2000 (n = 3) or 1000 ( n = 5)
```{r}
pink.long.red1 <- pink.long %>% # pink.long n = 72 obs
  filter(abundance < 2000) # pink.long.red1 n = 69 obs

pink.long.red2 <- pink.long %>% # pink.long.red2 n = 67 obs
  filter(abundance < 1000)

pink5.gam.red1 <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
                 data = pink.long.red1,
                 REML = F, family = nb())

pink5.gam.red2 <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"), 
                 data = pink.long.red2,
                 REML = F, family = nb())

par(mfrow=c(2,2))
gam.check(pink5.gam)
gam.check(pink5.gam.red1)
gam.check(pink5.gam.red2)
par(mfrow = c(1,1))

# need to load dharma more recently
library(DHARMa)
plot(DHARMa::simulateResiduals(pink5.gam))
plot(simulateResiduals(pink5.gam.red1))
plot(simulateResiduals(pink5.gam.red2))

```


# Final models w/ REML
```{r}
chum.final <- gam(abundance ~ Habitat + Mon + s(Lat1, bs = "tp") + s(SiteID, bs = "re"), data = chum.long,
                  family = nb(), method = "ML")

summary(chum.final)

# you can't use emmeans or glht out of the box with gams. 

pink.final <- gam(abundance ~ Habitat + Mon + s(SiteID, bs = "re"),
                      data = pink.long, method = "REML", family = nb())
summary(pink.final)

```

# Visualize Chum
using this: https://rdrr.io/cran/glmmTMB/man/predict.glmmTMB.html
```{r}
# what do we include for month
ggplot(chum.long) +
  geom_point(aes(x = Mon, y = abundance, color = Habitat)) # so the greatest catches for chum are in April and June
# lets use April for now
table(chum.long$Mon) # 135 seems reasonable, tahts mid may

par(mfrow=c(2,2))
visreg(chum.final)
visreg(chum.final, scale = "response")
par(mfrow=c(1,1))
summary(chum.final)

#chum.predicts <- tidygam::predict_gam(chum.final, tran_fun = exp) # need to do this to exp log-odds of the negative binomial
#plot(chum.predicts, series = "Habitat")

# okay so for predicting random effects models likely you want to predict to the "average" site - so you might not include it in the predict fucntion. 
# chum.final includes re of siteID, habitat and month
pr.habc <- predict(chum.final, 
                   newdata = data.frame(Habitat = unique(chum.long$Habitat), 
                                        Mon = "May", 
                                        Lat1 = rep(median(chum.long$Lat1, 
                                                          na.rm = T),2),
                                        SiteID = NA),
                   na.action = NULL,
                   exclude = "s(SiteID)",
                   se.fit = TRUE)

pr.monc <- predict(chum.final, 
                   newdata = data.frame(Habitat = "Eelgrass", 
                                        Mon = unique(chum.long$Mon), 
                                        Lat1 = rep(median(chum.long$Lat1, 
                                                          na.rm = T), 4),
                                        SiteID = rep(NA, 4)),
                   na.action = NULL,
                   exclude = "s(SiteID)",
                   se.fit = TRUE)

newdat.lat <- data.frame(Habitat = "Eelgrass", Mon = "May",
                         Lat1 = seq(min(chum.long$Lat1),
                                    max(chum.long$Lat1),
                                    length.out = 100),
                         SiteID = NA)
pr.latc <- predict(chum.final, newdata = newdat.lat, se.fit = TRUE, na.action = NULL, exclude = "s(SiteID)")

pred <- exp(pr.habc$fit) 
lwr <- exp(pr.habc$fit - (1.96*pr.habc$se.fit))
upr <- exp(pr.habc$fit + (1.96*pr.habc$se.fit))

pred.m <- exp(pr.monc$fit)
lwr.m <- exp(pr.monc$fit - (1.96*pr.monc$se.fit))
upr.m <- exp(pr.monc$fit + (1.96*pr.monc$se.fit))

newdat.lat$fit <- exp(pr.latc$fit)
newdat.lat$lwr <- exp(pr.latc$fit - (1.96*pr.latc$se.fit))
newdat.lat$upr <- exp(pr.latc$fit + (1.96*pr.latc$se.fit))

chum.long$Habitat <- as.character(chum.long$Habitat)

pr.habc.df <- data.frame(cbind(pred, lwr, upr, unique(chum.long$Habitat))) %>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(pred),
         lwr = as.numeric(lwr),
         upr = as.numeric(upr))

chum.long$Mon <- as.character(chum.long$Mon)

pr.mon.df <- data.frame(cbind(pred.m, lwr.m, upr.m, unique(chum.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(pred.m),
         lwr = as.numeric(lwr.m),
         upr = as.numeric(upr.m))
```

## Chum graph (Fig 2)
```{r}
chum.long$Mon<- plyr::revalue(chum.long$Mon, c("Apr"="April", "May"="May", "Jun"= "June", "Jul"="July")) 
                           
# make sure when you're putting the graph together that you include lwr and upr without the fit cause those are the lwr and upr confidence intervals. 
habc <- ggplot(pr.habc.df) +
  geom_col(aes(x = Habitat, y = fit, fill = Habitat), fill = "lightgray") + 
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = chum.long, aes(x = Habitat, y = abundance)) +
  plot_theme() +
  theme(legend.position="none") +
  labs(y = NULL) 
habc


# now month
monc <- pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "lightgray") + 
  geom_errorbar(aes(x = Mon, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = chum.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  plot_theme() +
  labs(x = "Month", y = NULL) 

monc

# now latitude
latc <- ggplot(newdat.lat) +
  geom_ribbon(aes(x = Lat1, ymax = upr, ymin = lwr), fill = "lightgray") +
  geom_line(aes(x = Lat1, y = fit), color = "black") +
  geom_point(data = chum.long, aes(x = Lat1, y = abundance)) +
  #scale_y_continuous(trans = "log10") #+
  plot_theme() +
  labs(x = "Latitude", y = NULL) 


gam.chum <- habc + monc + latc + plot_spacer() +
  plot_annotation(tag_levels="a") + 
  plot_layout(nrow = 2)

gam.chum.plots <- add_global_label(gam.chum, 
                                   Ylab = "Juvenile~Chum~Salmon~(no.~seine^-1)",
                 size = 8.25)


#ggsave("Manuscript_scripts&data/Figures&Tables/Obj1_chum_gam_plot.png", plot = gam.chum.plots, width = 14, height = 10)
```

# Visualize Pink
```{r}
# lets use May for now
table(pink.long$Mon) # 135 seems reasonable
table(pink.long$Mon, pink.long$abundance>0) # more non-zero catches in may compared to any other month... could this be way the CI doesn't go as low? 

visreg(pink.final)
visreg(pink.final, scale = "response")
summary(pink.final)
# okay so for predicting random effects models likely you want to predict to the "average" site - so you might not include it in the predict fucntion. 
# chum.final includes re of siteID, habitat and month
pr.habp <- predict(pink.final, 
                   newdata = data.frame(Habitat = unique(pink.long$Habitat), 
                                        Mon = "May", 
                                        SiteID = NA), se.fit = TRUE, na.action = NULL)

pr.monp <- predict(pink.final, 
                   newdata = data.frame(Habitat = "Eelgrass", 
                                        Mon = unique(pink.long$Mon), 
                                        SiteID = NA), se.fit = TRUE, na.action = NULL)


# if 
#log_normal_correct_p<- (sd(residuals(pink.final))^2)/2


predp <- exp(pr.habp$fit)
lwrp <- exp(pr.habp$fit - (1.96*pr.habp$se.fit))
uprp <- exp(pr.habp$fit + (1.96*pr.habp$se.fit))

predp.m <- exp(pr.monp$fit)
lwrp.m <- exp(pr.monp$fit - (1.96*pr.monp$se.fit))
uprp.m <- exp(pr.monp$fit + (1.96*pr.monp$se.fit))

pink.long$Habitat <- as.character(pink.long$Habitat)

pr.habp.df <- data.frame(cbind(predp, lwrp, uprp, unique(pink.long$Habitat)))%>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(predp),
         lwr = as.numeric(lwrp),
         upr = as.numeric(uprp))

pink.long$Mon <- as.character(pink.long$Mon)

pr.monp.df <- data.frame(cbind(predp.m, lwrp.m, uprp.m, unique(pink.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(predp.m),
         lwr = as.numeric(lwrp.m),
         upr = as.numeric(uprp.m))
```

## Pink graph (Fig 3)
```{r}
pink.long$Mon <- plyr::revalue(pink.long$Mon, c("Apr"="April", "May"="May", "Jun"= "June", "Jul"="July"))
                           

habp <- ggplot(pr.habp.df) +
  geom_col(aes(x = Habitat, y = fit, fill = Habitat)) + #, fill = "lightgray") +
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = pink.long, aes(x = Habitat, y = abundance)) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans = "log10") +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
   #                  breaks = c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000)) +
  scale_fill_manual(values = c("#339900", "#D16103")) +
  theme(legend.position="none") +
  labs(y = Juvenile~Pink~Salmon~(no.~seine^-1))
habp


# now doy
monp <- pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "#7c7c7c") + # fill = "lightgray"
  geom_errorbar(aes(x = Mon, ymax = upr, ymin = lwr), width = 0.2) +
  geom_jitter(data = pink.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  #ylim(c(0, 8000)) +
  plot_theme() +
  #ylim(c(0, 9000)) +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
   #                  breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) +
  labs(x = "Month", y = NULL) 
monp

glmm.pink.plot <- habp + monp + plot_annotation(tag_levels="a")
glmm.pink.plot.colored <- monp + habp

p_ranges_y <- c(ggplot_build(glmm.pink.plot[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(glmm.pink.plot[[2]])$layout$panel_scales_y[[1]]$range$range)

glmm.pink.plot2 <- glmm.pink.plot &
  ylim(min(p_ranges_y), max(p_ranges_y)) &
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) 


p_ranges_y2 <- c(ggplot_build(glmm.pink.plot.colored[[1]])$layout$panel_scales_y[[1]]$range$range,
                ggplot_build(glmm.pink.plot.colored[[2]])$layout$panel_scales_y[[1]]$range$range)

glmm.pink.plot.colored2 <- glmm.pink.plot.colored &
  ylim(min(p_ranges_y2), max(p_ranges_y2)) &
  scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),
                     breaks= (c(0, 1, 10, 100, 1000, 2000, 3000, 6000, 9000))) 

#ggsave("chp1_obj1_glmm.pink.plt_logscale.colored.png", plot = glmm.pink.plot.colored2, width = 16, height = 7)

#ggsave("chp1_obj1_glmm.pink.plot_logscale.png", plot = glmm.pink.plot2, width = 14, height = 8)
```

Non-exponentiated CI of the predicted pink.final
```{r}
pr.habp <- predict(pink.final, newdata = data.frame(Habitat = unique(pink.long$Habitat), Mon = "May", SiteID = NA), se.fit = TRUE)

pr.monp <- predict(pink.final, newdata = data.frame(Habitat = "Eelgrass", Mon = unique(pink.long$Mon), SiteID = NA), se.fit = TRUE)

predp <- pr.habp$fit
lwrp <- pr.habp$fit - (1.96*pr.habp$se.fit)
uprp <- pr.habp$fit + (1.96*pr.habp$se.fit)

predp.m <- pr.monp$fit
lwrp.m <- pr.monp$fit - (1.96*pr.monp$se.fit)
uprp.m <- pr.monp$fit + (1.96*pr.monp$se.fit)

pink.long$Habitat <- as.character(pink.long$Habitat)

pr.habp.df <- data.frame(cbind(predp, lwrp, uprp, unique(pink.long$Habitat)))%>%
  dplyr::rename(Habitat = V4) %>%
  mutate(fit = as.numeric(predp),
         lwr = as.numeric(lwrp),
         upr = as.numeric(uprp))

pink.long$Mon <- as.character(pink.long$Mon)

pr.monp.df <- data.frame(cbind(predp.m, lwrp.m, uprp.m, unique(pink.long$Mon))) %>%
  dplyr::rename(Mon = V4) %>%
  mutate(fit = as.numeric(predp.m),
         lwr = as.numeric(lwrp.m),
         upr = as.numeric(uprp.m))

ggplot(pr.habp.df) +
  geom_col(data = pr.habp.df, aes(x = Habitat, y = fit, fill = Habitat)) + #fill = "lightgray") +
  geom_errorbar(aes(x = Habitat, ymax = upr, ymin = lwr), width = 0.2, data = pr.habp.df) #+
  #geom_jitter(data = pink.long, aes(x = Habitat, y = log(abundance+1))) +
  #ylim(c(-0.1, 7)) +
  plot_theme() +
  #scale_y_continuous(trans = "log10") +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10)) +
  #scale_fill_manual(values = c("#339900", "#D16103")) +
  theme(legend.position="none") +
  labs(y = Juvenile~Pink~Salmon~(no.~seine^-1))

pr.mon.df %>%
  mutate(Mon = ifelse(Mon == "Apr", "April", Mon),
         Mon = ifelse(Mon == "Jun", "June", Mon),
         Mon = ifelse(Mon == "Jul", "July", Mon)) %>%
  ggplot() +
  geom_col(aes(x = Mon, y = fit), fill = "#7c7c7c") +
  geom_errorbar(aes(x = Mon, ymax = fit + upr, ymin = fit - lwr), width = 0.2) +
  #geom_jitter(data = pink.long, aes(x = Mon, y = abundance)) +
  scale_x_discrete(limits = c("April", "May", "June", "July")) +
  #ylim(c(0, 8000)) +
  plot_theme() +
  #scale_y_continuous(trans=scales::pseudo_log_trans(base = 10)) +
  #ylim(c(0, 10000)) +
  labs(x = "Month", y = NULL) 
```


#  Model selection table 
## Chum model selection (Supp Table 2)
```{r}
cand.models <- list() # create an empty list to include all the candidate models
cand.models[[1]] <- ch.re1 # add in  models one at a time
cand.models[[2]] <- ch.re2

cand.models2 <- list()
cand.models2[[1]] <- chum1
cand.models2[[2]] <- chum.zinf

cand.models3 <- list()
cand.models3[[1]] <- chum1.gam
cand.models3[[2]] <- chum2.gam
cand.models3[[3]] <- chum3.gam
cand.models3[[4]] <- chum5.gam

Modnames.re <- data.frame(names = c(paste(c(formula(ch.re1), formula(ch.re2)))),
                           method = c("REML", "REML"),
                           family = c("Neg. Binomial", "Neg. Binomial"))

Modnames.fix <- data.frame(names = c(paste(c(formula(chum1), formula(chum.zinf)))),
                           method = c("ML", "ML"),
                           family = c("Neg. Binomial", "Zero-inflated Neg. Binomial"))

Modnames3 <- data.frame(names = c(paste(c(formula(chum1.gam), formula(chum2.gam),
                                          formula(chum3.gam),
                                          formula(chum5.gam)))),
                        method = rep("ML", 4))

# create dataframe with model names. 
cand.models

d <- data.frame()
for(i in 1:length(cand.models)) {
  mod <- cand.models[[i]]
  name <- paste(c(formula(mod)))
  #aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, 
                     AICc = aicc, BIC = bic)
  d <- rbind(d, temp)
}

# now for the fx effects model
d2 <- data.frame()
for(i in 1:length(cand.models2)) {
  mod <- cand.models2[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d2 <- rbind(d2, temp)
}

# fixed effects w/ spatial term
d3 <- data.frame()
for(i in 1:length(cand.models3)) {
  mod <- cand.models3[[i]]
  name <- paste(c(formula(mod)))
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, 
                     AICc = aicc, BIC = bic)
  d3 <- rbind(d3, temp)
}

d.chum.re <- d %>%
  left_join(Modnames.re, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.chum.fx <- d2 %>%
  mutate(family = c("Neg. Binomial", "Zero-inflated Neg. Binomial")) %>%
  left_join(Modnames.fix, by = c("model.names" = "names", "family")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.chum3 <- d3 %>%
  mutate(family = rep("Neg. Binomial", 4)) %>%
  left_join(Modnames3, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  arrange(AICc)

d.chum <- rbind(d.chum.re, d.chum3)
# best model includes random effect of site and habitat and month as fixed using negative binomial not zero inflated

#all.models <- as.data.frame(Modnames.all)
#write.csv(d.chum, "Manuscript_scripts&data/Figures&Tables/Obj1_Chum_candidate.csv")
```

## Pink model selection (Supp Table 3)
```{r}
cand.models <- list() # create an empty list to include all the candidate models
cand.models[[1]] <- p.re1 # add in  models one at a time
cand.models[[2]] <- p.re2

cand.models2 <- list()
cand.models2[[1]] <- pink1
cand.models2[[2]] <- pink.zinf

cand.models3 <- list()
cand.models3[[1]] <- pink1.gam
cand.models3[[2]] <- pink2.gam
cand.models3[[3]] <- pink3.gam
cand.models3[[4]] <- pink4.gam
cand.models3[[5]] <- pink5.gam


Modnames.re <- data.frame(names = c(paste(c(formula(p.re1), formula(p.re2)))),
                           method = c("REML", "REML"))

Modnames.fix <- data.frame(names = c(paste(c(formula(pink1), formula(pink.zinf)))),
                           method = c("ML", "ML"),
                           family = c("Neg. Binomial", "Zero-inflated Neg. Binomial"))

Modnames3 <- data.frame(names = c(paste(c(formula(pink1.gam), formula(pink2.gam),
                                          formula(pink3.gam), formula(pink4.gam),
                                          formula(pink5.gam)))),
                        method = rep("ML", 5))

# create dataframe with model names. 
cand.models

d <- data.frame()
for(i in 1:length(cand.models)) {
  mod <- cand.models[[i]]
  name <- paste(c(formula(mod)))
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, 
                     AICc = aicc, BIC = bic)
  d <- rbind(d, temp)
}

# now for the fx effects model
d2 <- data.frame()
for(i in 1:length(cand.models2)) {
  mod <- cand.models2[[i]]
  name <- paste(c(formula(mod)))
  aic <- AIC(mod)
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, AIC = aic, 
                     AICc = aicc, BIC = bic)
  d2 <- rbind(d2, temp)
}

# fixed effects w/ spatial term
d3 <- data.frame()
for(i in 1:length(cand.models3)) {
  mod <- cand.models3[[i]]
  name <- paste(c(formula(mod)))
  aicc <- AICc(mod)
  bic <- BIC(mod)
  #df <- mod$df.null # this doesn't work with re models
  sums <- summary(mod)
  #dev.expl <- sums$dev.expl*100 # also not in re models
  temp <- data.frame(model.names = name, 
                     AICc = aicc, BIC = bic)
  d3 <- rbind(d3, temp)
}

d.pink.re <- d %>%
  left_join(Modnames.re, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.pink.fx <- d2 %>%
  mutate(family = c("Neg. Binomial", "Zero-inflated Neg. Binomial")) %>%
  left_join(Modnames.fix, by = c("model.names" = "names", "family")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AIC = AIC - min(AIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2)

d.pink3 <- d3 %>%
  left_join(Modnames3, by = c("model.names" = "names")) %>%
  arrange(BIC) %>%
  mutate(delta.BIC = BIC - min(BIC),
         delta.AICc = AICc - min(AICc)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  arrange(AICc)


d.pink <- rbind(d.pink.re, d.pink3)
#write.csv(d.pink, "Manuscript_scripts&data/Figures&Tables/Obj1_Pink_candidateTable.csv")
```

best model from table
```{r}
d.chum
summary(chum.final)
```

best model from table pink
```{r}
d.pink
summary(pink.final)
```


# Best model parameter estimates (Table 1)
```{r}
library(sjPlot)
summary(pink.final)
summary(chum.final)
table1 <- tab_model(chum.final, pink.final,
          transform = NULL, # transform NULL is impt
          dv.labels = c("Juvenile Chum Salmon", "Juvenile Pink Salmon"),
          show.est = TRUE, string.est = "Estimate", show.r2 = FALSE,
          show.stat = TRUE, string.stat = "Statistic",
          p.threshold = c(0.1, 0.05, 0.01, 0.001),
          show.se = TRUE, show.ci = FALSE, string.p = "P-Value",
          string.se = "Standard Error", p.style = "numeric")
           
          
          
          # pred.labels = c("Intercept", "Habitat [kelp]","Non-linear date", 
          #                  "Site (random)", "Date"))

```

